\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}

\newcommand{\E}{\mathbb{E}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\OPT}{\operatorname{OPT}}
\newcommand{\optpost}{\operatorname{OPT}_{\mathrm{post}}}
\newcommand{\optadapt}{\operatorname{OPT}_{\mathrm{adapt}}}

\title{A Recursive Triangle-like Asymmetric Instance with Clairvoyance Gap Tending to $4/3$}
\author{}
\date{}

\begin{document}
\maketitle

\section{The construction}

We define a family of asymmetric stochastic TSP instances.  Distances are taken to be the directed shortest-path metric induced by the directed edges described below.  The depot is always present.  The client vertices are independent, with probabilities $1/2$.

For every integer $l\geq 0$, define a directed recursive block $B_l$.  Each block has a set of permanent vertices, denoted $P_l$.  These permanent vertices will be deterministic.  For $l=0$, the block $B_0$ consists of a single permanent vertex.

For $l\geq 1$, the block $B_l$ is obtained from two disjoint copies of $B_{l-1}$, denoted $B_{l-1}^{\mathrm L}$ and $B_{l-1}^{\mathrm R}$, together with a new midpoint vertex $m_l$.  

Let $w_l := 2^{l-1}$, then for every permanent vertex $u \in P(B_{l-1}^{\mathrm L})$ and every permanent vertex $v\in P(B_{l-1}^{\mathrm R})$, add a directed edge
$u \to v$ of length $w_l$.
For every $v \in P(B_{l-1}^{\mathrm R})$, add a directed edge $v\to m_l $ of length $w_l$.  
For every $u \in P(B_{l-1}^{\mathrm L})$, add a directed edge $m_l \to u$ of length $w_l$. 
Thus, after contracting the two child blocks, the top level of $B_l$ is the directed triangle
\[
    B_{l-1}^{\mathrm L}
    \longrightarrow
    B_{l-1}^{\mathrm R}
    \longrightarrow
    m_l
    \longrightarrow
    B_{l-1}^{\mathrm L},
\]
where each displayed arc has length $w_l$.

The permanent set of $B_l$ is
\[
    P_l := P(B_{l-1}^{\mathrm L})\cup P(B_{l-1}^{\mathrm R}).
\]
Thus $|P_l|=2^l$.  All permanent vertices are active with probability $1$.  Every midpoint vertex is active independently with probability $1/2$.

The full rooted instance $I_l$ is obtained by adding a depot $r$.  For every permanent vertex $u \in P_l$, add directed edges
\[
    r \to u,
    \qquad
    u \to r,
\]
each of length $w_l=2^{l-1}$.

\section{Statement}

\begin{theorem}\label{thm:recursive-gap}
For every $l\geq 1$, the instance $I_l$ satisfies
\[
    \optpost(I_l) \le (3l+4)2^{l-2},
\]
and
\[
    \optadapt(I_l) \ge (l+1)2^l.
\]
Consequently,
\[
    \frac{\optadapt(I_l)}{\optpost(I_l)}
    =
    \frac{4(l+1)}{3l+4}
    \longrightarrow
    \frac43.
\]
\end{theorem}

The rest of the note proves the theorem.

\section{A posteriori cost}

We explicitly give an a posteriori solution. The solution works in a recursive way. Since a posteriori algorithm knows
certain vertices would be active or not, we can make recursive decisions based on this.

For a block $B_l$, if the midpoint $m_l$ is active, then we first recursively traverse $B_{l-1}^R$, then we visit $m_l$, finally we 
recursively traverse $B_{l-1}^L$; if the midpoint $m_l$ is inactive, then we first traverse $B_{l-1}^L$, then we traverse $B_{l-1}^R$.
If we denote the cost of such recursive algorithm traversing $B_l$, excluding the cost of the first step goes from root $r$ and the last step
goes back to $r$, to be $T_l$, then we have the recurrence equation:

\begin{align*}
    T_l = 2 T_{l-1} + \frac{3}{2} 2^{l-1} 
\end{align*}

Together with the initial condition that $T_1 = 3/2$, we have $T_l = \frac{3}{4} l 2^l$. Adding the cost of root, we have that for 
such a posteriori solution, the expected cost is $(\frac{3}{4}l+1) 2^l$. Thus we have:

\begin{align*}
   \optpost \le  (\frac{3}{4}l+1) 2^l
\end{align*}


\section{Adaptive cost}

We now analyze the optimal adaptive policy.  



\section{The ratio}

Combining the two formulas,
\[
    \frac{\optadapt(I_l)}{\optpost(I_l)}
    =
    \frac{(l+1)2^l}{(3l+4)2^{l-2}}
    =
    \frac{4(l+1)}{3l+4}.
\]
Therefore
\[
    \lim_{l\to\infty}
    \frac{\optadapt(I_l)}{\optpost(I_l)}
    =
    \frac43.
\]
For example, when $l=3$,
\[
    \optpost(I_3)=26,
    \qquad
    \optadapt(I_3)=32,
\]
and hence
\[
    \frac{\optadapt(I_3)}{\optpost(I_3)}
    =
    \frac{32}{26}
    =
    \frac{16}{13}
    \approx 1.23077.
\]
For $l=4$, the formulas give
\[
    \optpost(I_4)=64,
    \qquad
    \optadapt(I_4)=80,
\]
so the ratio is $5/4$.

\begin{remark}
In the implementation, the depot should be treated as always present.  Thus, if the list of activation probabilities includes the depot as vertex $0$, its probability should be $1$, not $1/2$.  The midpoint probabilities are $1/2$, and the permanent leaves have probability $1$.
\end{remark}

\end{document}
