%! TeX program = lualatex
\documentclass[12pt,a4paper,oneside]{report}

\input{preamble/usepackages}
\input{preamble/theoremenv}
\input{preamble/mathdefs}

\addbibresource{preamble/ref.bib}

\newenvironment{comments}
  {\par\medskip\begingroup\color{red!70!black}%
   \noindent\textbf{Comments:}%
   \begin{enumerate}[leftmargin=*,itemsep=0.25em,topsep=0.4em]}
  {\end{enumerate}\endgroup\medskip}

\newcommand{\sharat}[1]{\textcolor{red}{Sharat: #1}} %%% don't remove this macro

\begin{document}

% ============================================================
% FRONT MATTER
% ============================================================

\pagenumbering{Roman}
\input{preamble/title}
\maketitlepage
\pagenumbering{roman}

% --- Abstract ---
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

We study the stochastic traveling salesperson problem, where each client is active
independently according to a given probability and all active clients must be
served at minimum expected travel cost.

There are three common benchmarks that differ in when the active clients are revealed
and when the decisions are  made: an a~posteriori solution knows the complete
active set in advance and decide the tour after active clients are revealed,
an adaptive solution decide the next client to visit based on past information,
and an a~priori solution commits to a fixed tour ordering before any activations are known.

We compare the three benchmarks and give lower bounds for the gap between them.
For clients with symmetric distances, we construct instances with an adaptivity gap approaching~$5/4$ and
instances with a clairvoyance gap of~$11/10$. For asymmetric distances, we
construct a recursive family whose clairvoyance gap approaches~$4/3$. We also
give that, for every fixed $\varepsilon>0$, a deterministic
$(1{+}\varepsilon)$-approximation algorithm for a~priori TSP with running time
$c_\varepsilon^n\poly(n)$ based on dynamic programming.

\begin{comments}
  \item Too long. An abstract should convey the high level problem statement
  without getting into technical details. You have to accomplish this without
  introducing ambiguity.
  \item Rephrase. Sounds LLMie.
  \item You haven't even defined what symmetric instances are.
  \item Single stochastic vertex setting isn't an important case, especially
  because the associated result is not really exploiting the true structure.
  \item What's a directed metric? Also, does the distinction matter?
  \item It's important to mention that we use dynamic programming for this
  result but anything more is unnecessary. Nearby pairs introduces more
  confusion since the reader hasn't yet seen the cost expression for a~priori
  TSP.
  \item My general advice is to keep rewriting the abstract in your own words.
  Read it until it sounds like a coherent piece of text that someone in our
  field can understand.
\end{comments}
\tableofcontents

% ============================================================
% CHAPTER 1: INTRODUCTION
% ============================================================
\pagenumbering{arabic}
\chapter{Introduction}\label{ch:intro}

Many real-world routing and scheduling problems are subject to uncertainty.
In a delivery service, for example, a courier must plan a route through several customer locations, yet some customers may have no packages to receive or their requests may be canceled.
A classical abstraction is the \emph{stochastic traveling salesperson problem} (stochastic TSP), where each vertex~$v$ comes with a specified probability~$p_v$ and becomes \emph{active} (requires service) independently with probability~$p_v$.
The goal is to find a tour that visits all active vertices and minimizes the expected travel cost, where the cost comes from the underlying metric on vertices.

An important modeling choice is whether the salesperson is allowed to adapt future decisions based on information revealed during the execution of the tour.
In the \emph{nonadaptive} (a~priori) setting, the salesperson commits to a fixed tour before observing the realization; during execution the salesperson follows this tour while shortcutting inactive vertices.
In the \emph{adaptive} setting, decisions may depend on observations revealed along the way, allowing the tour to be adjusted on the fly.
In the \emph{a~posteriori} setting, the salesperson has full knowledge of the realization.

A fundamental question is how much adaptivity can improve performance relative to an optimal nonadaptive tour, and how much full clairvoyance improves over the best adaptive policy.
Note that computing an optimal adaptive policy (or an optimal a~priori tour) for stochastic TSP is computationally hard, since the setting with $p_v=1$ for all~$v$ corresponds to the classical TSP, which is known to be APX-hard~\cite{Karpinski2015}.
Also, describing an optimal adaptive policy might require exponential space.
Thus, we are interested in understanding the separations between the three benchmarks, and in designing policies or algorithms whose expected cost can be provably bounded.

\section{Our contributions}\label{sec:contributions}

We obtain the following results.

\begin{enumerate}
    \item \textbf{Adaptivity gap lower bound for symmetric metrics} (Chapter~\ref{ch:sym-adapt}).
    We construct a family of cycle-subdivided-clique instances and show that the ratio $\optprior/\optadapt$ tends to~$5/4$ as the number of vertices grows (Theorem~\ref{thm:cycle-subdivision}).
    Since $\optpost \le \optadapt$, the same lower bound applies to the obliviousness gap.

    \item \textbf{Clairvoyance gap lower bound for symmetric metrics} (Chapter~\ref{ch:sym-clair}).
    We exhibit a family of symmetric instances, the \emph{three-chain construction}, achieving $\optadapt/\optpost = 11/10$ for infinitely many~$n$
    (Theorem~\ref{thm:sym-clair}).
    For context, we also record an elementary upper bound of~$1/(1-p+p^2)$ for the restricted setting with a single stochastic vertex of activation probability~$p$ (Proposition~\ref{prop:general-p}).

    \item \textbf{Clairvoyance gap lower bound for asymmetric metrics} (Chapter~\ref{ch:asym-clair}).
    We define a recursive family of directed instances for which $\optadapt/\optpost$ tends to~$4/3$ (Theorem~\ref{thm:recursive-gap}).

    \item \textbf{Approximate a~priori TSP} (Chapter~\ref{ch:dp}).
    We give a $(1{+}\varepsilon)$-approximation algorithm for a~priori TSP on directed metrics, running in time $c_\varepsilon^n\poly(n)$
    (Theorem~\ref{thm:directed-apriori-ptas}).
    The algorithm discards pairs for which the probability that all intervening clients are inactive is tiny, losing only a negligible amount, and then uses a dynamic program whose state stores just the relevant suffix of the partial order.
\end{enumerate}

\section{Related work}\label{sec:related}

\paragraph{Classical TSP.}
The traveling salesperson problem is one of the most studied problems in combinatorial optimization.
For symmetric metrics, the classical $3/2$-approximation of Christofides~\cite{Christofides1973} stood for decades until Karlin, Klein, and Oveis~Gharan~\cite{Karlin2021} obtained a slightly improved factor.
For asymmetric metrics, Svensson, Tarnawski, and V\'egh~\cite{Svensson2020} gave the first constant-factor approximation.

\paragraph{A~priori TSP.}
The a~priori model was introduced by Jaillet~\cite{Jaillet1985}.
Shmoys and Talwar~\cite{Shmoys2008} proved that the obliviousness gap $\optprior/\optpost$ is at most~$3$ for symmetric metrics, yielding a constant-factor approximation via a master tour.
Schalekamp and Shmoys~\cite{SchalekampShmoys2008} studied the connection between universal and a~priori TSP.
Blauth, Neuwohner, Puhlmann, and Vygen~\cite{Blauth2024} recently obtained improved approximation guarantees for a~priori TSP.
For asymmetric metrics, Christalla, Puhlmann, and Traub~\cite{Christalla2025} showed that the obliviousness gap is $O(\sqrt{n})$ and $\Omega(n^{1/4}/\log n)$; moreover, their construction also implies an adaptivity gap of $\Omega(n^{1/4}/\log n)$.

\paragraph{Adaptive and stochastic TSP.}
Gupta, Nagarajan, and Ravi~\cite{GuptaNagarajanRavi2017} studied adaptive TSP and gave an $O(\log^2 n)$-approximation for the stochastic version.
Ene, Nagarajan, and Saket~\cite{EneNagarajanSaket2017} studied approximation algorithms for stochastic $k$-TSP.
Jiang, Li, Liu, and Singla~\cite{JiangLiLiuSingla2020} obtained improved algorithms and adaptivity gaps for stochastic $k$-TSP.
Fu, Li, and Xu~\cite{FLX18} gave a PTAS for a class of stochastic dynamic programs.

\paragraph{Adaptivity gaps in stochastic optimization.}
The study of adaptivity gaps---how much adaptivity helps in stochastic optimization---has a rich history beyond TSP.
Dean, Goemans, and Vondr\'ak~\cite{DeanGoemansVondrak2008} studied the benefit of adaptivity for stochastic knapsack.
Gupta and Nagarajan~\cite{GuptaNagarajan2013} introduced the stochastic probing framework and proved constant adaptivity gaps for matching.
Gupta, Nagarajan, and Singla~\cite{GuptaNagarajanSingla2016,GuptaNagarajanSingla2017} extended these results to submodular and XOS objectives.
Brada\v{c}, Singla, and \v{Z}u\v{z}i\'c~\cite{BradacSinglaZuzic2019} obtained near-optimal adaptivity gaps for stochastic multi-value probing.
Levin and Vainer~\cite{LV14} studied adaptivity in the stochastic blackjack knapsack problem.
Barak and Talgam-Cohen~\cite{BarakTalgamCohen2026} recently studied semi-adaptivity gaps for stochastic knapsack.

% ============================================================
% MODEL AND BENCHMARKS
% ============================================================
\section{Model and benchmarks}\label{sec:model}

\subsection{Metric spaces and tours}

Let $(\{r\}\cup V,d)$ be a finite metric space on $n+1$ vertices (a depot vertex~$r$ and $n$~clients) equipped with a distance function~$d$.
That is, $d$ satisfies the triangle inequality: for any three vertices $u,v,w$, $d(u,v) \leq d(u,w) + d(w,v)$.
A metric is \emph{symmetric} if $d(x,y)=d(y,x)$ for all $x,y$; otherwise it
may be asymmetric, in which case distances are directed.

A \emph{tour} on a subset $S \subseteq \{r\}\cup V$ containing the depot is a cyclic sequence $T=(r,u_1,u_2,\ldots,u_{|S|-1},r)$ that visits every vertex in~$S$ exactly once. Its \emph{cost} is
\[
\cost(T) := d(r,u_1) + d(u_1,u_2) + \cdots + d(u_{|S|-2},u_{|S|-1}) + d(u_{|S|-1},r).
\]
For a subset $A\subseteq V$, we write $\opt(A)$ for the minimum cost of a tour on $\{r\}\cup A$.

\subsection{Three benchmarks for stochastic TSP}

In the stochastic setting, the set of active vertices is a subset $A \subseteq V$ generated by independent activations: each vertex $v \in V$ is active (i.e., $v \in A$) independently with probability~$p_v$. We denote by~$q$ the resulting product distribution on~$2^V$, so $q(A) = \prod_{v \in A}p_v \cdot \prod_{v \notin A}(1 - p_v)$.

We consider three benchmarks for stochastic TSP, distinguished by how much the solution may depend on the realization of the active set. In each case we define the cost of a feasible solution via an overloaded function $\cost(\cdot)$.

\begin{enumerate}
    \item \textbf{A posteriori TSP.}
        An \emph{a~posteriori solution} is a family $\mathcal{S}=(T_A)_{A\subseteq V}$, where $T_A$ is a tour on $\{r\}\cup A$ for each activation set~$A$.
        We define
        \[
        \cost(\mathcal{S}) := \E_{A\sim q}\bigl[\cost(T_A)\bigr].
        \]
        Since each $T_A$ can be chosen independently, the optimum is
        \[
        \optpost := \min_{\mathcal{S}}\,\cost(\mathcal{S}) = \E_{A\sim q}\bigl[\opt(A)\bigr].
        \]

    \item \textbf{A priori (nonadaptive) TSP.}
        An \emph{a~priori tour} is a tour $T=(r,v_1,\ldots,v_n,r)$ on the full vertex set $\{r\}\cup V$.
        For a realization~$A$, the \emph{shortcut} $T\vert_A$ is obtained by retaining only the vertices in $\{r\}\cup A$ in the cyclic order prescribed by~$T$.
        More precisely, if $1 \leq i_1 < i_2 < \cdots < i_{|A|} \leq n$ are such that $\{v_{i_k} : 1 \leq k \leq |A|\} = A$, then $T\vert_A = (r,v_{i_1},v_{i_2},\ldots,v_{i_{|A|}},r)$.
        We overload:
        \[
        \cost(T) := \E_{A\sim q}\bigl[\cost(T\vert_A)\bigr].
        \]
        The optimum is $\optprior := \min_{T}\,\cost(T)$, where the minimum is over all a~priori tours.

    \item \textbf{Adaptive TSP.}
        The salesperson knows all the activation probabilities but not the realization~$A$.
        During execution, the salesperson calls clients one by one:
        if the called client~$v$ is in~$A$, the salesperson \emph{must} travel to~$v$ immediately; otherwise the salesperson remains in place.
        The order of calls may depend on the outcomes observed so far.

        Formally, an adaptive policy is a function $\pi: (\{r\} \cup V) \times 2^V \rightarrow \{r\} \cup V \cup \{\bot\}$, where $\pi(u,S)$ denotes the next vertex to call when the salesperson is at~$u$ and the set of uncalled vertices is~$S$.
        We require: (a)~$\pi(u,S) := \bot$ whenever $u \in S$ (invalid input); (b)~$\pi(u,S) = r$ iff $S = \emptyset$ (return to depot only when done); (c)~$\pi(u,S) \in S$ iff $S \neq \emptyset$ (must call an uncalled vertex).

        Given a realization~$A$, the policy induces a random tour $T(A;\pi)$ on $\{r\}\cup A$ as follows.
        Initialize $u \leftarrow r$, $S \leftarrow V$.
        While $S \neq \emptyset$: compute $v = \pi(u,S)$; if $v \in A$ then travel to~$v$ and set $u\leftarrow v$; set $S \leftarrow S \setminus \{v\}$.
        Finally, return to~$r$.
        We overload:
        \[
        \cost(\pi) := \E_{A\sim q}\bigl[\cost(T(A;\pi))\bigr].
        \]
        The optimum is $\optadapt := \min_{\pi}\,\cost(\pi)$, where the minimum is over all adaptive policies.
\end{enumerate}

\subsection{Gap measures}\label{sec:gaps}

Among the three benchmarks, a~posteriori TSP has the most information (the full realization~$A$), a~priori TSP has the least (only the metric and probabilities), and adaptive TSP lies in between.

We define an \emph{instance} $\mathcal{I} := (V \cup \{r\}, d, \{p_v\})$ of stochastic TSP. For any instance~$\mathcal{I}$,
\begin{equation}\label{eq:optineq}
\optpost \le \optadapt \le \optprior
\end{equation}
since from left to right the solution has less information about the realization, resulting in higher optimal cost.

We define the following gap measures:
\begin{itemize}
    \item \emph{Adaptivity gap}: $\sup_{\mathcal{I}} \f{\optprior}{\optadapt}$ --- the worst-case benefit of adaptivity over a fixed a~priori tour.
    \item \emph{Clairvoyance gap}: $\sup_{\mathcal{I}} \f{\optadapt}{\optpost}$ --- the benefit of full clairvoyance (knowing $A$ upfront) over the best adaptive policy.
    \item \emph{Obliviousness gap}: $\sup_{\mathcal{I}} \f{\optprior}{\optpost}$ --- combines both effects.
\end{itemize}
We remark that the literature on a~priori TSP uses the term adaptivity gap for what we call the obliviousness gap; we deviate from this convention following \cite{Christalla2025} for the sake of clarity.

For symmetric metrics, the obliviousness gap is known to be at most $O(1)$; in particular, $\optprior/\optpost \le 3$ is shown in~\cite{Shmoys2008}.
The same upper bound carries over to the clairvoyance gap and the adaptivity gap by~\eqref{eq:optineq}.
For asymmetric metrics, Christalla, Puhlmann, and Traub~\cite{Christalla2025} show that the obliviousness gap is $O(\sqrt{n})$ and $\Omega(n^{1/4} \cdot \log^{-1}n)$.
Moreover, the construction in Section~3.1 of~\cite{Christalla2025} also implies an adaptivity gap of $\Omega(n^{1/4} \cdot \log^{-1}n)$.

\begin{comments}
  \item Preliminaries section is usually meant to state well known tools and
  techniques that will be used later in the thesis. For instance, we will be
  using some linear programming based lower bounds. Some results on
  circulations, Hoffmann's condition etc. Some concentration inequalities. The
  current content should be moved to the introduction.
\end{comments}

\section{Organization}

Chapter~\ref{ch:sym-adapt} proves the adaptivity gap lower bound of~$5/4$ for symmetric metrics.
Chapter~\ref{ch:sym-clair} proves the symmetric clairvoyance-gap lower bound of~$11/10$ and records an auxiliary single-stochastic-vertex upper bound.
Chapter~\ref{ch:asym-clair} constructs the asymmetric recursive instance with clairvoyance gap tending to~$4/3$.
Chapter~\ref{ch:dp} presents the $(1{+}\varepsilon)$-approximation algorithm for directed a~priori TSP.
Chapter~\ref{ch:conclusion} concludes with a summary and directions for future
research.
Appendix~\ref{app:evidence} records exploratory computational searches for the symmetric clairvoyance gap.


% ============================================================
% CHAPTER 2: ADAPTIVITY GAP — SYMMETRIC
% ============================================================
\chapter{Adaptivity Gap for Symmetric Metrics}\label{ch:sym-adapt}

Our main result in this chapter is a lower bound of $5/4$ on the adaptivity gap for symmetric stochastic TSP. We show that there exists a family of instances, obtained by subdividing edges in a Hamiltonian cycle of a complete graph, for which the ratio $\optprior/\optadapt$ tends to $5/4$ as the number of vertices goes to infinity. 

\begin{theorem}\label{thm:cycle-subdivision}
The adaptivity gap for symmetric stochastic TSP is at least $\frac54$.
\end{theorem}

\begin{corollary}\label{cor:cycle-oblivious}
The obliviousness gap for symmetric stochastic TSP is at least~$\frac54$.
\end{corollary}

\begin{proof}
This follows from Theorem~\ref{thm:cycle-subdivision} and
$\optpost\le\optadapt$.
\end{proof}

We prove Theorem~\ref{thm:cycle-subdivision} by exhibiting a family of instances and establishing a matching pair of bounds. Consider the following construction.

\begin{definition}[Cycle-subdivided clique]\label{def:cycle-sub-graph}
Fix $n\ge 7$, and let $v_1,\dots,v_n$ be the vertices of the complete graph
$K_n$, with indices taken modulo~$n$. For every $i\in[n]$, subdivide the edge
$v_iv_{i+1}$ by introducing a new vertex~$v_i'$. Denote the resulting
unweighted graph by~$G_n$.
\end{definition}

The \emph{cycle-subdivided-clique instance} $\mathcal I_n$ has depot $r:=v_1$ and client set
\[
V_n:=\{v_2,\dots,v_n\}\cup\{v_1',\dots,v_n'\}.
\]
The underlying distance function is the shortest-path metric induced by~$G_n$. Every
original client $v_2,\dots,v_n$ is active with probability~$1$, while every
subdivision vertex~$v_i'$ is active independently with probability~$1/2$.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    scale=1.05,
    original/.style={rectangle,draw=black,fill=white,minimum size=8mm,inner sep=0pt,font=\small},
    stochastic/.style={circle,draw=black,fill=gray!20,minimum size=7mm,inner sep=0pt,font=\small},
    cycleedge/.style={black,line width=0.6pt},
    chord/.style={black,line width=0.6pt}
]
\foreach \i in {1,...,6} {
  \coordinate (c\i) at ({90-60*(\i-1)}:3.2);
}
\foreach \i in {1,...,6} {
  \pgfmathtruncatemacro{\next}{mod(\i,6)+1}
  \coordinate (m\i) at ($(c\i)!0.5!(c\next)$);
}
\foreach \i in {1,...,6} {
  \foreach \j in {1,...,6} {
    \ifnum\j>\i
      \pgfmathtruncatemacro{\next}{mod(\i,6)+1}
      \ifnum\j=\next\else
        \ifnum\i=1\ifnum\j=6\else\draw[chord] (c\i)--(c\j);\fi
        \else\draw[chord] (c\i)--(c\j);\fi
      \fi
    \fi
  }
}
\foreach \i in {1,...,6} {
  \pgfmathtruncatemacro{\next}{mod(\i,6)+1}
  \draw[cycleedge] (c\i)--(m\i)--(c\next);
}
\foreach \i in {1,...,6} {
  \node[original] (v\i) at (c\i) {$v_\i$};
  \node[stochastic] (vp\i) at (m\i) {$v_\i'$};
}
\node[above=2mm of v1,font=\small] {depot $r$};
\end{tikzpicture}
\caption{The cycle-subdivided clique $G_6$. Original vertices are active with
probability~$1$ and are represented as squares; subdivision vertices are active
with probability~$1/2$ and are represented as gray circles.}
\label{fig:cycle-subdivision}
\end{figure}

Theorem~\ref{thm:cycle-subdivision} follows from the next two lemmas.

\begin{lemma}\label{lem:cycle-prior}
For every $n\ge 7$, the instance $\mathcal I_n$ satisfies $\optprior=\frac{15}{8}n$.
\end{lemma}

\begin{lemma}\label{lem:cycle-adapt}
For every $n\ge 7$, the instance $\mathcal I_n$ satisfies $\optadapt\le \frac32 n+6$.
\end{lemma}

\begin{proof}[Proof of Theorem~\ref{thm:cycle-subdivision}]
Lemma~\ref{lem:cycle-prior} gives $\optprior=\frac{15}{8}n$
and Lemma~\ref{lem:cycle-adapt} gives $\optadapt\le \frac32 n+6$.
Hence $\optprior/\optadapt$ converges to $\frac{5}{4}$ as $n$ grows.
\end{proof}

\sharat{It's good practice to collect all tools that you need for a technical proof before you start the proof. I would rephrase the below sentence as follows: 
The rest of this chapter is devoted to proving Lemmas blah and blah. 
Start a new section entitled a priori optimum for cycle-subdivided clique instances. The next section would be adaptive optimum for ...}

\sharat{In the a prior section, you start with a short paragraph on how the proof works. It's good practice to give the reader an idea on where the proof is going before you take them down the rabbit hole. Starting a new section gives you a chance to collect properties of the metric without entering the proof environment.}

We now prove the two lemmas in turn. %% this will be commented out. 

\begin{proof}[Proof of Lemma~\ref{lem:cycle-prior}]
Fix an arbitrary a~priori tour $T=(v_1,s_1,s_2,\dots,s_{2n-1},v_1)$ on $\{v_1\}\cup V_n$, where $s_1,\dots,s_{2n-1}$ is a permutation of $\{v_2,\dots,v_n,v_1',\dots,v_n'\}$. We shall prove that its expected realized length is at least $15n/8$, and then exhibit a tour attaining this value.

\sharat{It's not standard practice in our field to introduce subheadings in proofs like metric structure and decomposition into segments. These things are better isolated even before you start the proof.}

\medskip
\noindent\emph{Metric structure.}
For later use we record the relevant distances in the metric completion.
For original vertices,
\[
d(v_i,v_{i+1})=2,
\qquad
 d(v_i,v_j)=1 \quad \text{if } j\not\equiv i\pm 1 \pmod n.
\]
For subdividing vertices,
\[
d(v_i',v_i)=d(v_i',v_{i+1})=1,
\qquad
 d(v_i',v_j)=2 \quad \text{if } j\notin\{i,i+1\}.
\]
Finally, for $i\neq j$,
\[
d(v_i',v_j')=
\begin{cases}
2, & \text{if the cycle edges } v_iv_{i+1} \text{ and } v_jv_{j+1} \text{ are adjacent},\\
3, & \text{otherwise.}
\end{cases}
\]

\sharat{The metric structure is just an observation that should be collected early on (outside the proof) with labeled equations. Give each of the 3 settings a unique equation number.}

\medskip
\noindent\emph{Decomposition into segments.}
Deleting all subdividing vertices from $T$ leaves a cyclic ordering
\[
(v_1=u_1,u_2,\dots,u_n,v_1)
\]
of the original vertices $v_1,\dots,v_n$. For each $j\in[n]$, let
\[
W_j=(w_{j,1},\dots,w_{j,k_j})
\]
be the ordered list (possibly empty) of subdividing vertices appearing on $T$ between $u_j$ and $u_{j+1}$, where $u_{n+1}:=v_1$. Then
\[
\sum_{j=1}^n k_j=n,
\]
since every subdividing vertex appears in exactly one such segment.

Given original vertices $u,v$ and an ordered sequence $W=(w_1,\dots,w_k)$ of distinct subdividing vertices, write
\[
\Phi(u;W;v)
\]
for the expected length of the realized path obtained from the segment
\[
(u,w_1,\dots,w_k,v)
\]
after deleting inactive subdividing vertices. Because all original vertices are always active, the realized tour decomposes into these segment contributions, and therefore
\[
\E[\cost(T|_A)] = \sum_{j=1}^n \Phi(u_j;W_j;u_{j+1}).
\]
Thus it is enough to derive a lower bound on $\Phi(u;W;v)$ in terms of $k:=|W|$.

\sharat{Ideally, the proof of Lemma A should end here. Claims 2.6 and long segment analysis are the technical heart of the proof which are distracting the reader from knowing what is happening in the proof of Lemma A. In the proof sketch/overview paragraph at the beginning of this section, you should already allude to the strategy briefly. The proof of lemma is dependent on the reader believing that $f(k) >= 1 + 7k/8$. This should be made explicit and then the proof should be concluded. Proving the technical bits is a separate proof.}

\medskip
\noindent\emph{Short segments.}
We first treat the cases $k\le 3$.

\begin{claim}\label{clm:short-seg}
For every segment $(u;W;v)$ with $k=|W|\le 3$,
\[
\Phi(u;W;v)\ge
\begin{cases}
1, & k=0,\\
2, & k=1,\\
\dfrac{11}{4}, & k=2,\\
\dfrac{29}{8}, & k=3.
\end{cases}
\]
Moreover, equality is attained for $k=2$ by $(v_i,v_i',v_{i+1}',v_{i+1})$, and for $k=3$ by $(v_{i+1},v_{i+1}',v_i',v_{i-1}',v_i)$.
\end{claim}

\sharat{11/4 and 29/8 are overlapping. Just remove the dfrac macro.}

\begin{proof}
If $k=0$, then the segment contributes the single edge $uv$, so $\Phi(u;\varnothing;v)=d(u,v)\ge 1$.

If $k=1$, say $W=(w)$, then
\[
\Phi(u;w;v)=\tfrac12 d(u,v)+\tfrac12\bigl(d(u,w)+d(w,v)\bigr).
\]
If $u$ and $v$ are consecutive on the original cycle, then $d(u,v)=2$ and also $d(u,w)+d(w,v)\ge 2$. If they are not consecutive, then $d(u,v)=1$, and no subdividing vertex is incident with both $u$ and $v$, so $d(u,w)+d(w,v)\ge 3$. In either case $\Phi(u;w;v)\ge 2$.

Now let $k=2$, say $W=(w_1,w_2)$. There are four realizations.
If $d(u,v)=1$, then neither $w_1$ nor $w_2$ is incident with both $u$ and $v$, so each singleton realization has cost at least~$3$, while the realization with both active has cost at least~$4$. Hence
\[
\Phi(u;W;v)\ge \frac{1+3+3+4}{4}=\frac{11}{4}.
\]
If $d(u,v)=2$, then exactly one subdividing vertex is incident with both $u$ and~$v$, so at most one singleton realization has cost~$2$; the two singleton costs therefore sum to at least~$5$, and the realization with both active has cost at least~$4$. Thus
\[
\Phi(u;W;v)\ge \frac{2+5+4}{4}=\frac{11}{4}.
\]
Equality is attained, for example, by $(v_i,v_i',v_{i+1}',v_{i+1})$, whose four realization costs are $2,2,3,4$.

Finally let $k=3$, say $W=(w_1,w_2,w_3)$.
Suppose first that $d(u,v)=1$. Then each singleton realization has cost at least~$3$. A two-vertex realization can have cost~$4$ only if one active subdivider is incident with~$u$, the other is incident with~$v$, and the corresponding cycle edges are adjacent. For fixed nonconsecutive $u$ and $v$, there is at most one unordered pair of subdividing vertices with this property. Hence among the three two-vertex realizations, at most one has cost~$4$, and the other two have cost at least~$5$. The realization with all three active has cost at least~$6$. Therefore
\[
\Phi(u;W;v)\ge \frac{1+9+14+6}{8}=\frac{15}{4}>\frac{29}{8}.
\]

Now suppose that $d(u,v)=2$. Then exactly one subdividing vertex is incident with both $u$ and~$v$, so among the three singleton realizations at most one has cost~$2$. Hence the sum of the singleton costs is at least $2+3+3=8$. A two-vertex realization can have cost~$4$ only if one active subdivider is the unique subdivider of the edge $uv$ and the other comes from one of the two cycle edges adjacent to $uv$. Thus among the three two-vertex realizations, at most two have cost~$4$, and the remaining one has cost at least~$5$; the sum of these three costs is therefore at least~$13$. The realization with all three active has cost at least~$6$. Consequently,
\[
\Phi(u;W;v)\ge \frac{2+8+13+6}{8}=\frac{29}{8}.
\]
Equality is attained, for example, by $(v_{i+1},v_{i+1}',v_i',v_{i-1}',v_i)$, whose eight realization costs are
\[
2,\ 3,\ 2,\ 3,\ 4,\ 5,\ 4,\ 6.
\]
\end{proof}

\medskip
\noindent\emph{Long segments.}
We now consider a segment
\[
(u;w_1,\dots,w_k;v)
\]
with $k\ge 4$. Set
\[
z_0:=u,\qquad z_t:=w_t\ (1\le t\le k),\qquad z_{k+1}:=v.
\]
Two retained vertices $z_i,z_j$ become consecutive after shortcutting precisely when every vertex strictly between them is deleted. Since each $w_t$ is retained independently with probability~$1/2$, we obtain the exact expansion
\begin{align*}
\Phi(u;W;v)
&=2^{-k}d(u,v)
 +\sum_{t=1}^k 2^{-t}d(u,w_t)
 +\sum_{t=1}^k 2^{-(k-t+1)}d(w_t,v) \\
&\qquad
 +\sum_{1\le i<j\le k}2^{-(j-i+1)}d(w_i,w_j).
\end{align*}
Introduce the indicators
\[
\alpha_t:=\1_{\{w_t \text{ is incident with } u\}},
\qquad
\beta_t:=\1_{\{w_t \text{ is incident with } v\}},
\]
and, for $1\le i<j\le k$,
\[
\gamma_{ij}:=\1_{\{w_i,w_j \text{ correspond to adjacent cycle edges}\}}.
\]
By the distance formulas above,
\begin{align*}
d(u,v)&=1+\1_{\{u,v \text{ are consecutive on the original cycle}\}},\\
d(u,w_t)&=2-\alpha_t,\\
d(w_t,v)&=2-\beta_t,\\
d(w_i,w_j)&=3-\gamma_{ij}.
\end{align*}
Substituting these identities into the preceding formula yields
\[
\Phi(u;W;v)=1+\frac{3k}{2}+2^{-k}\1_{\{u,v\text{ consecutive}\}}-D,
\]
where
\[
D:=\sum_{t=1}^k 2^{-t}\alpha_t
 +\sum_{t=1}^k 2^{-(k-t+1)}\beta_t
 +\sum_{1\le i<j\le k}2^{-(j-i+1)}\gamma_{ij}.
\]
We now estimate~$D$.
Since an original vertex is incident with exactly two subdividing vertices,
\[
\sum_{t=1}^k 2^{-t}\alpha_t\le 2^{-1}+2^{-2}=\frac34,
\qquad
\sum_{t=1}^k 2^{-(k-t+1)}\beta_t\le \frac34.
\]
For the last sum, consider the graph on vertex set $\{w_1,\dots,w_k\}$ in which two vertices are adjacent when the corresponding subdividing vertices arise from adjacent edges of the original cycle. This graph is a subgraph of a cycle, hence has maximum degree at most~$2$ and therefore at most $k$ edges. Every edge contributes weight at most~$1/4$, so
\[
\sum_{1\le i<j\le k}2^{-(j-i+1)}\gamma_{ij}\le \frac{k}{4}.
\]
Therefore
\[
D\le \frac34+\frac34+\frac{k}{4}=\frac32+\frac{k}{4},
\]
and hence
\[
\Phi(u;W;v)\ge 1+\frac{3k}{2}-\left(\frac32+\frac{k}{4}\right)=\frac54 k-\frac12.
\]
Since $k\ge 4$,
\[
\frac54 k-\frac12\ge 1+\frac{7k}{8},
\]
so every segment with $k\ge 4$ satisfies
\[
\Phi(u;W;v)\ge 1+\frac{7k}{8}.
\]
Combining this with the previous claim, we conclude that every segment with $k$ subdividing vertices satisfies
\[
\Phi(u;W;v)\ge
\begin{cases}
1, & k=0,\\
1+\dfrac{7k}{8}, & k\ge 1.
\end{cases}
\]

\medskip
\noindent\emph{Global lower bound.}
Applying the segment bound to the decomposition of~$T$, we obtain
\begin{align*}
\E[\cost(T|_A)]
&=\sum_{j=1}^n \Phi(u_j;W_j;u_{j+1})\\
&\ge \sum_{j:k_j=0} 1+\sum_{j:k_j\ge 1}\left(1+\frac{7k_j}{8}\right)\\
&=n+\frac78\sum_{j=1}^n k_j\\
&=n+\frac78 n = \frac{15}{8}n.
\end{align*}
Since $T$ was arbitrary, this shows that
\[
\optprior\ge \frac{15}{8}n.
\]

\sharat{At a minimum, you should make these headings bold and not just use italics. Italics are in general used for inline writing where you want to emphasize something. They can be mistaken for a new sentence. ``Matching constructions.'' is not a proper sentence in this context.}
\medskip
\noindent\emph{Matching constructions.}

It remains to exhibit an a~priori tour, for each $n \geq 7$, attaining this value. 

\smallskip
\noindent\emph{Even $n$.}
Write $n=2m$. Consider the tour
\begin{align*}
T^{\mathrm{even}}
:=(&v_1,v_1',v_2',v_2,
   v_4,v_4',v_3',v_3,
   v_6,v_6',v_5',v_5,
   \dots,\\
  &v_{2m},v_{2m}',v_{2m-1}',v_{2m-1},v_1).
\end{align*}
This tour has $m$ nonempty segments, namely
\[
(v_1,v_1',v_2',v_2)
\quad\text{and}\quad
(v_{2t},v_{2t}',v_{2t-1}',v_{2t-1})\qquad (2\le t\le m),
\]
each with expected contribution $11/4$, together with $m$ direct edges between original vertices, each of length~$1$. Hence
\[
\E[\cost(T^{\mathrm{even}}|_A)] = m\left(\frac{11}{4}+1\right)=\frac{15}{8}n.
\]

\smallskip
\noindent\emph{Odd $n$.}
Write $n=2m+1$ with $m\ge 2$. Consider the tour
\begin{align*}
T^{\mathrm{odd}}
:=(&v_n,v_2,v_2',v_1',v_n',v_1,
   v_4,v_4',v_3',v_3,
   v_6,v_6',v_5',v_5,
   \dots,\\
  &v_{2m},v_{2m}',v_{2m-1}',v_{2m-1},v_n).
\end{align*}
This tour has one $3$-segment,
\[
(v_2,v_2',v_1',v_n',v_1),
\]
with expected contribution $29/8$, together with $m-1$ segments
\[
(v_{2t},v_{2t}',v_{2t-1}',v_{2t-1})\qquad (2\le t\le m),
\]
each contributing $11/4$, and $m+1$ direct edges between original vertices, each of length~$1$. Therefore
\begin{align*}
\E[\cost(T^{\mathrm{odd}}|_A)]
&=\frac{29}{8}+(m-1)\cdot \frac{11}{4}+(m+1)\\
&=\frac{15}{8}(2m+1)=\frac{15}{8}n.
\end{align*}

In both parity cases we have constructed an a~priori tour of expected cost $15n/8$. Combined with the lower bound, this proves
\[
\optprior=\frac{15}{8}n.
\]
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:cycle-adapt}]
We exhibit an adaptive policy~$\pi$. Let
\[
P \;:=\; v_2,\, v_2',\, v_3,\, \dots,\, v_{n-1}',\, v_n.
\]
\sharat{We should adopt proper terminology to talk about adaptive policies. In the introduction, calling terminology makes sense. That's how we motivate the problem. Somewhere in the introduction you will be formalizing the setup. At this point, the probing terminology should be introduced. Make it clear in the introduction that the probes are done remotely. Do include a prompt in your LLM to highlight all places where the terminology is inconsistent.}

First call $v_1'$ and $v_n'$ (the only clients outside~$P$), then return to the depot $r=v_1$.
In the worst case both are active, costing $d(r,v_1')+d(v_1',v_n')+d(v_n',r)=1+2+1=4$.
Next, move from $r$ to $v_2$, which costs $2$,  
% (cost~$2$) 
and process~$P$. \sharat{fix throughout.}
At each step the uncalled clients on~$P$ form a contiguous subpath, and the salesperson sits at one of its two endpoint originals. The policy probes the adjacent subdividing vertex: if it is active, traverse it and the next original (cost~$2$) \sharat{original vertex you mean} and continue in the same direction; if it is inactive, jump to the opposite endpoint original \sharat{???} of the remaining subpath (cost~$1$, since nonconsecutive originals are at distance~$1$) and proceed symmetrically. 
\sharat{incorrect use of parantheses. nonconsecutive originals is not good terminology. Opposite of original is duplicate. You should come up with a better naming system here.}

\sharat{Avoid using $a$ and $A$ as much as possible since the thesis is going to be filled with a priori and a posteriori everywhere.}

Let $a:=|A\cap\{v_2',\dots,v_{n-1}'\}|$. The $a$ active subdividers on~$P$ each contribute cost~$2$. Each of the $(n-2)-a$ inactive ones triggers a switch of cost~$1$, except possibly the last, which may cost~$2$. Including the return to~$r$ (cost at most~$2$), the middle phase costs at most $2+2a+((n-2)-a+1)+2=n+a+3$.
The total cost per realization is therefore at most $4+(n+a+3)=n+a+7$, and taking expectation with $\E[a]=(n-2)/2$ gives $\E[\cost(\pi)]\le \frac32 n+6$.
\end{proof}

\clearpage
\begin{comments}
  \item The second sentence should be put in a separate corollary.
  \item Edge length being~$1$ can be stated at the very end. The focus should
  be on defining the underlying graph.
  \item Since we control the indexing, there is no need to mention that we are
  choosing the trivial Hamiltonian cycle. Just say that each edge of the form
  $v_iv_{i+1}$ is subdivided by introducing a new vertex~$v_i'$. Don't use
  ``replace'' terminology.
  \item In fact, remove the wording of length from the definition. This
  shortest-path metric is the correct way of describing the instance.
  \item The stochastic TSP instance should be different from the
  cycle-subdivided graph. Introduce the graph first. While talking about the
  TSP instance, you say that the metric is obtained by the shortest path metric
  induced by the cycle-subdivided graph with vertices having so and so
  activation probabilities.
  \item Include a figure of the cycle-subdivided complete graph say for $n=6$.
  Make sure to draw vertices as boxes/circles as opposed to tiny dots since
  edges will be crossing at many places and we don't want the reader to think
  that there are a lot more vertices.
  \item Obliviousness gap is irrelevant here. It's just a corollary that
  shouldn't be talked about here.
  \item Replace $r$ by~$v_1$?
  \item Change cycle-subdivision graph to cycle-subdivided clique.
  \item No need to refer to this theorem as cycle-subdivision gap. It conflicts
  with the adaptive gap. It's an adaptivity gap construction arising from a
  metric induced by a cycle-subdivided clique.
  \item No need to say that line crossings are not vertices since the current
  representation of the nodes removes that ambiguity. Also, maybe use squares
  for permanent/original vertices.
  \item Put the corollary right after the statement of Theorem~2.1.
\end{comments}


% ============================================================
% CHAPTER 3: CLAIRVOYANCE GAP — SYMMETRIC
% ============================================================
\chapter{Clairvoyance Gap for Symmetric Metrics}\label{ch:sym-clair}

In this chapter we study the symmetric clairvoyance gap $\optadapt/\optpost$.
We construct an explicit family of symmetric instances achieving ratio~$11/10$.
We also record an auxiliary upper bound for the restricted setting in which
only one vertex is stochastic.

\section{The three-chain lower bound}\label{sec:3chain}

\begin{theorem}[Symmetric clairvoyance gap lower bound]\label{thm:sym-clair}
The clairvoyance gap for symmetric stochastic TSP is at least~$11/10$.
\end{theorem}

We prove the theorem by constructing a family of three-chain instances whose
adaptive and a~posteriori optimum values differ by precisely this factor.

\begin{construction}[Three-chain graph $H_k$]\label{con:3chain}
Let $k \ge 3$ be an integer.
Define a symmetric weighted graph with vertex set
\[
V(H_k) \;=\; A\;\cup\; B\;\cup\; C\;\cup\; \{s\},
\qquad |V(H_k)| = 3k+1,
\]
where the \emph{chains}
\begin{align*}
A &= (a_0, a_1, \ldots, a_{k-1}), &
B &= (b_0, b_1, \ldots, b_{k-1}), &
C &= (c_0, c_1, \ldots, c_{k-1})
\end{align*}
are vertex-disjoint paths. There are two types of edges in this graph:

\medskip\noindent\textbf{Internal edges} (weight~$1$): for each chain $X \in \{A,B,C\}$
and each $0\le i\le k-2$, include the edge $\{x_i, x_{i+1}\}$ of weight~$1$.

\medskip\noindent\textbf{Inter-chain edges} (weight~$2(k-1)$):
\begin{equation*}\label{eq:interchain}
\begin{gathered}
\{a_0, b_{k-1}\},\quad
\{a_0, c_{k-1}\},\quad
\{b_0, a_{k-1}\},\quad
\{c_0, a_{k-1}\},\quad
\{c_0, b_{k-1}\},\quad
\{b_0, s\},\quad
\{c_{k-1}, s\}.
\end{gathered}
\end{equation*}

\end{construction}

The instance has depot $r=a_0$ and $3k$ client 
$V(H_k)\setminus\{r\}$. Its distance function is the shortest-path metric
induced by $H_k$. The client $s$ is active with probability~$1/2$, and
every other client is deterministic. We denote the deterministic clients
\[
D:=V(H_k)\setminus\{r,s\}
\]

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    x=1cm,y=1cm,
    deterministic/.style={rectangle,draw=black,fill=white,minimum width=8.5mm,
      minimum height=7mm,inner sep=0pt,font=\scriptsize},
    depot/.style={deterministic,fill=gray!15},
    stochastic/.style={circle,draw=black,fill=gray!30,minimum size=7.5mm,
      inner sep=0pt,font=\scriptsize},
    internal/.style={black,line width=0.75pt},
    crossing/.style={black!75,densely dashed,line width=0.75pt},
    rowlabel/.style={font=\small\bfseries,anchor=east}
]
% Chain A, drawn from a_0 to a_{k-1}.
\node[depot]         (A0) at (0, 2.4) {$a_0$};
\node[deterministic] (A1) at (1.5, 2.4) {$a_1$};
\node[deterministic] (Ak) at (6, 2.4) {$a_{k-1}$};

% Chain B is reversed so that its relevant endpoints align with A and C.
\node[deterministic] (Bk) at (0, 0) {$b_{k-1}$};
\node[deterministic] (B1) at (4.5, 0) {$b_1$};
\node[deterministic] (B0) at (6, 0) {$b_0$};

% Chain C, drawn from c_0 to c_{k-1}.
\node[deterministic] (C0) at (0,-2.4) {$c_0$};
\node[deterministic] (C1) at (1.5,-2.4) {$c_1$};
\node[deterministic] (Ck) at (6,-2.4) {$c_{k-1}$};
\node[stochastic]    (S)  at (7.3,-1.2) {$s$};

% The seven inter-chain edges, all of weight 2(k-1).
\draw[crossing] (A0)--(Bk);
\draw[crossing] (A0)--(Ck);
\draw[crossing] (B0)--(Ak);
\draw[crossing] (C0)--(Ak);
\draw[crossing] (C0)--(Bk);
\draw[crossing] (B0)--(S);
\draw[crossing] (Ck)--(S);

% Unit-weight edges within the three chains.
\draw[internal] (A0)--(A1)--(Ak);
\draw[internal] (Bk)--(B1)--(B0);
\draw[internal] (C0)--(C1)--(Ck);

% Ellipses cover the long middle segments of the schematic chains.
\node[fill=white,inner sep=3pt] at (3.45, 2.4) {$\cdots$};
\node[fill=white,inner sep=3pt] at (2.1, 0) {$\cdots$};
\node[fill=white,inner sep=3pt] at (3.45,-2.4) {$\cdots$};

\node[rowlabel] at (-0.65, 2.4) {chain $A$};
\node[rowlabel] at (-0.65, 0) {chain $B$};
\node[rowlabel] at (-0.65,-2.4) {chain $C$};
\node[above=1mm of A0,font=\scriptsize] {depot $r$};
\end{tikzpicture}
\caption{The three-chain graph $H_k$. Each horizontal row is a chain of
edges with weight $1$; chain~$B$ is drawn in reverse order for simplicity. The dashed edges
have weight~$2(k-1)$. The shaded square is the depot, and the gray circle is the
only stochastic client.}
\label{fig:three-chain}
\end{figure}

\sharat{Replace weight by length throughout.}

The inter-chain weight $2(k-1)$ balances the cost of a crossing against
the cost of traversing the chains and makes the active and inactive optimal
tour skeletons compete at the thresholds used below.

Theorem \ref{thm:sym-clair} can be directly derived from the following lemmas:

\begin{lemma}[Optimal a posteriori cost]\label{lem:sym-apost-cost}
    The three-chain instance $H_k$ satisfies $\optpost = 10(k-1)$.
\end{lemma}
\begin{lemma}[Optimal adaptive cost]\label{lem:sym-adapt-cost}
    The three-chain instance $H_k$ satisfies $\optadapt = 11(k-1)$.
\end{lemma}
\begin{proof}[Proof to Theorem~\ref{thm:sym-clair}]
    Follows immediately from definition.
\end{proof}

The proof of these identities has two structural ingredients.  First we identify the only
near-optimal crossing patterns in the two realizations.  We then show that an
adaptive policy cannot realize both patterns because its calls before learning
the status of~$s$ must be the same in both realizations.

\begin{lemma}[Realization costs and near-optimal routes]\label{lem:near-optimal-routes}
Let $k\ge3$ and $h=k-1$.
In the instance $G_k$, the following hold.
\begin{enumerate}
\item[\textup{(i)}]
Every tour on $\{r\}\cup D$ has cost at least $9h$.
This bound is attained by
\begin{multline*}
T_D:\;
a_0 \to a_1 \to \cdots \to a_{k-1}
\xrightarrow{\;2h\;} b_0 \to b_1 \to \cdots \to b_{k-1}\\
\xrightarrow{\;2h\;} c_0 \to c_1 \to \cdots \to c_{k-1}
\xrightarrow{\;2h\;} a_0,
\end{multline*}
which has cost $9h$.
Moreover, any tour on $\{r\}\cup D$ of cost strictly less than $11h$
has the following property: every expansion of its metric edges into shortest
paths in $H_k$ uses each of these three inter-chain crossings exactly once
and uses no other crossing:
\[
    \{a_0,c_{k-1}\},\qquad
    \{c_0,b_{k-1}\},\qquad
    \{b_0,a_{k-1}\}.
\]

\item[\textup{(ii)}]
Every tour on $\{r\}\cup D\cup\{s\}$ has cost at least $11h$.
This bound is attained by
\begin{multline*}
T_{\mathrm{all}}:\;
a_0 \to a_1 \to \cdots \to a_{k-1}
\xrightarrow{\;2h\;} c_0 \to c_1 \to \cdots \to c_{k-1}\\
\xrightarrow{\;2h\;} s
\xrightarrow{\;2h\;} b_0 \to b_1 \to \cdots \to b_{k-1}
\xrightarrow{\;2h\;} a_0,
\end{multline*}
which has cost $11h$.
Moreover, any tour on the full vertex set of cost strictly less than $13h$
has the following property: every shortest-path expansion uses each of these
four crossings exactly once and uses no other crossing:
\[
    \{a_0,b_{k-1}\},\qquad
    \{b_0,s\},\qquad
    \{s,c_{k-1}\},\qquad
    \{c_0,a_{k-1}\}.
\]
\end{enumerate}
\end{lemma}

\begin{proof}
Replace every metric edge of a tour by a shortest path in the underlying
weighted graph.  The resulting closed walk has the same length, and its
support---the subgraph formed by the edges it uses---is connected.

We first show that the internal edges of each chain contribute at least~$h$.
Fix one chain.  If all its $h$ internal edges occur in the walk, the claim is
immediate.  Otherwise, at most one internal edge can be absent: two absent
edges would separate the required vertices between them from the support of
the walk, since only the two endpoints of the chain have edges leaving the
chain.  If one internal edge is absent, each of the remaining $h-1$ internal
edges is a bridge in the support and is therefore traversed at least twice by
the closed walk.  Its internal-edge contribution is then at least
$2(h-1)\ge h$, where the last inequality uses $h\ge2$.  Summing over the three
chains gives a contribution of at least~$3h$.

Call a traversal of one of the seven edges in~\eqref{eq:interchain} a
\emph{crossing}.  Each crossing has cost~$2h$.
For a tour on $\{r\}\cup D$, the expanded walk starts and ends in chain~$A$
and must visit chains~$B$ and~$C$, so it uses at least three crossings.  Hence
its cost is at least $3h+6h=9h$.
For a tour on $\{r\}\cup D\cup\{s\}$, the expanded walk must also visit~$s$,
so it uses at least four crossings.  Hence its cost is at least
$3h+8h=11h$.

It remains to identify the tours that can be close to these lower bounds.
First consider a tour on $\{r\}\cup D$ of cost strictly less than
$11h=3h+8h$. It must use exactly three crossings.  After contracting each
chain, these crossings form the cycle $A$--$B$--$C$--$A$; in particular, the
walk does not use~$s$.  The $B$--$C$ crossing is necessarily
$\{b_{k-1},c_0\}$.  There are two choices for the $A$--$B$ crossing and two
for the $A$--$C$ crossing.  Choosing $\{a_0,b_{k-1}\}$ together with
$\{a_0,c_{k-1}\}$ repeats the crossing endpoint on chains $A$ and~$B$;
choosing $\{a_0,b_{k-1}\}$ together with $\{a_{k-1},c_0\}$ repeats it on
$B$ and~$C$; and choosing $\{a_{k-1},b_0\}$ together with
$\{a_{k-1},c_0\}$ repeats it on $A$ and~$C$.  The remaining combination is
the unique one using distinct endpoints on every chain, and its crossing set
is
\[
E_D:=\bigl\{\{a_0,c_{k-1}\},\{c_0,b_{k-1}\},
              \{b_0,a_{k-1}\}\bigr\}.
\]
If the two crossings incident with a chain use the same endpoint, the walk
must travel from that endpoint to the other end of the chain and back, at
cost at least~$2h$.  Hence each of the other three combinations pays at least
$2h+2h+h=5h$ on chain edges and at least $5h+6h=11h$ in total, a
contradiction.  Thus the three crossings of every tour below~$11h$ are
exactly those in~$E_D$.

Now consider a full tour of cost strictly less than
$13h=3h+10h$.  It uses exactly four crossings, which form a cycle through
$A,B,s,C$.  The two crossings incident with~$s$ must be $\{b_0,s\}$ and
$\{c_{k-1},s\}$.  Again there are two choices for the $A$--$B$ crossing and
two for the $A$--$C$ crossing.  The three choices other than
$\{a_0,b_{k-1}\}$ together with $\{a_{k-1},c_0\}$ are as follows:
$\{a_0,b_{k-1}\}$ with $\{a_0,c_{k-1}\}$ repeats endpoints on $A$ and~$C$;
$\{a_{k-1},b_0\}$ with $\{a_{k-1},c_0\}$ repeats them on $A$ and~$B$; and
$\{a_{k-1},b_0\}$ with $\{a_0,c_{k-1}\}$ repeats them on $B$ and~$C$.
Hence the unique combination using distinct endpoints on every chain is
\[
E_{\mathrm{all}}:=
\bigl\{\{a_0,b_{k-1}\},\{b_0,s\},\{s,c_{k-1}\},
        \{c_0,a_{k-1}\}\bigr\}.
\]
Each of the other three combinations repeats an endpoint on two chains, so
the same $5h$ chain-edge lower bound gives total cost at least
$5h+8h=13h$.  Thus every full tour below~$13h$ uses exactly the four
crossings in~$E_{\mathrm{all}}$.
\end{proof}

\begin{lemma}[Incompatibility of the two near-optimal skeletons]
\label{lem:incompatible-skeletons}
Let $C_0$ and $C_1$ be the realized costs of an adaptive policy when $s$ is
inactive and active, respectively. Then it is impossible to have both
$C_0<11h$ and $C_1<13h$.
\end{lemma}

\begin{proof}
The two realizations make exactly the same calls until the policy calls~$s$,
because every other client is deterministic. Write their common called prefix
as
\[
P=(r=x_0,x_1,\dots,x_t=u).
\]
Thus the inactive realized tour begins with~$P$, whereas the active realized
tour begins with $(P,s)$.

Assume for contradiction that $C_0<11h$ and $C_1<13h$. Expand every metric
edge into a shortest path in $H_k$, choosing the same expansion for the
edges of the common prefix. Lemma~\ref{lem:near-optimal-routes} forces the
inactive and active expanded tours to use the crossing sets $E_D$ and
$E_{\mathrm{all}}$, respectively. These sets are disjoint. Consequently, the
common prefix contains no crossing. Since it starts at $a_0$, every client
called before~$s$ lies on chain~$A$, and $u\in A$.

Only $b_0$ and $c_{k-1}$ are adjacent to~$s$ in the underlying graph. The
expanded path from $u$ to~$s$ in the active realization must therefore reach
$s$ through one of them. Because its crossings must belong to
$E_{\mathrm{all}}$, there are only two possibilities:
\begin{itemize}
  \item it goes from $A$ through $\{a_0,b_{k-1}\}$, traverses chain~$B$ from
  $b_{k-1}$ to~$b_0$, and then uses $\{b_0,s\}$; or
  \item it goes from $A$ through $\{a_{k-1},c_0\}$, traverses chain~$C$ from
  $c_0$ to~$c_{k-1}$, and then uses $\{c_{k-1},s\}$.
\end{itemize}
In the first case no client of~$B$ has yet been called. Passing through its
vertices in the shortest-path expansion does not call them, so all of~$B$
must still be served after reaching~$s$. But the two crossings just used are
the only crossings of $E_{\mathrm{all}}$ incident with~$B$; the remaining
skeleton proceeds from $s$ through $C$ and $A$ and never returns to~$B$.
Serving~$B$ would therefore require repeating a crossing or adding another
one, contradicting the forced four-crossing skeleton. The second case gives
the same contradiction with chain~$C$. Hence the two inequalities cannot
hold simultaneously.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:sym-clair}]
Since $s$ is active with probability~$1/2$, there are exactly two
realizations.  Lemma~\ref{lem:near-optimal-routes} gives optimum costs $9h$
when $s$ is inactive and $11h$ when it is active.  Therefore
\[
    \optpost=\frac12\cdot9h+\frac12\cdot11h=10h.
\]

For an adaptive upper bound, follow the order of $T_{\mathrm{all}}$ and
call~$s$ after $c_{k-1}$.  If $s$ is active, continue through~$s$ and then
chain~$B$.  If it is inactive, call $b_0$ next; the metric edge from
$c_{k-1}$ to~$b_0$ costs at most $4h$, by the path
$c_{k-1}$--$s$--$b_0$.  Thus either realization costs at most
$\cost(T_{\mathrm{all}})=11h$, and $\optadapt\le11h$.

For the reverse inequality, consider any adaptive policy and let $C_0$ and
$C_1$ be its realized costs when $s$ is inactive and active.  If
$(C_0+C_1)/2<11h$, the realization lower bounds $C_0\ge9h$ and $C_1\ge11h$
force $C_0<11h$ and $C_1<13h$.  This contradicts
Lemma~\ref{lem:incompatible-skeletons}.  Hence $\optadapt\ge11h$.

Substituting $h=k-1$ gives the displayed optimum values above.  Their ratio is
$11/10$, which proves the claimed lower bound.
\end{proof}

\begin{remark}[Tour restructuring]\label{rem:restructuring}
The two optimal realized tours visit the chains in opposite orders:
$T_{\mathrm{all}}$ routes $A \to C \to s \to B$,
while $T_D$ routes $A \to B \to C$.
The cost reduction from removing~$s$ is not a local vertex deletion
but a \emph{global tour restructuring} that saves one inter-chain crossing.
\end{remark}

\begin{remark}[Equality of adaptive and a~priori]
The order $T_{\mathrm{all}}$ is also an a~priori tour of expected cost
$11(k-1)$ or less, since shortcutting an inactive~$s$ replaces the two edges
$c_{k-1}s$ and $sb_0$ by a metric edge of no greater length.  Therefore
$\optprior\le 11(k-1)$.  Together with
$\optadapt\le\optprior$ from~\eqref{eq:optineq}, this gives
$\optprior=\optadapt=11(k-1)$ for this family.  All of the gap comes from
clairvoyance, not from adaptivity.
\end{remark}

\section{An auxiliary single-stochastic-vertex bound}\label{sec:single-vertex-upper}

\sharat{You don't have to call it an auxiliary bound. It's mostly a thing we say while talking but not necessarily write it.}

This section records an elementary bound for the restricted case of a single
stochastic vertex. It is independent of the three-chain argument and is not
claimed to be tight. Consider a symmetric instance with a single stochastic
vertex~$s$ of activation probability $p\in(0,1)$ and deterministic
customers~$D$.
If $D=\emptyset$, the adaptive and a~posteriori benchmarks coincide, so for
the notation and proof below we may assume $D\ne\emptyset$.
Write $T_{\mathrm{all}}$ for the optimal tour on $D\cup\{s\}\cup\{r\}$ and $T_D$ for the optimal tour on $D\cup\{r\}$, with costs
\[
    \optall:=\cost(T_{\mathrm{all}}),
    \qquad
    \opt_D:=\cost(T_D).
\]
Let
\[
    \rho:=\frac{\optall}{\opt_D},
    \qquad
    d_{\min}:=\min_{v\in D\cup\{r\}}d(v,s),
    \qquad
    \beta:=\frac{d_{\min}}{\opt_D}.
\]

\begin{proposition}[Single-stochastic-vertex bound]\label{prop:general-p}
For any symmetric metric instance with a single stochastic vertex~$s$
of probability~$p \in (0,1)$,
\[
\frac{\optadapt}{\optpost} \;\le\; \frac{1}{1 - p + p^2}.
\]
The right-hand side is maximized at $p = 1/2$, where it equals~$4/3$;
we do not claim that this value is attained by an instance.
By continuity, at $p = 0$ and $p = 1$ the bound equals~$1$ (no gap).
\end{proposition}

\begin{proof}
For general~$p$, the a~posteriori expected cost is
\[
\optpost = (1-p)\,\opt_D + p\,\optall = \opt_D\bigl((1-p) + p\rho\bigr).
\]
One adaptive policy simply follows $T_{\mathrm{all}}$.  Another follows
$T_D$ to a closest vertex $v$ to~$s$ and
calls~$s$ there.  If $s$ is inactive, it continues along $T_D$.  If $s$ is
active, it calls the next vertex of $T_D$ after visiting~$s$; by the triangle
inequality, this changes the next leg by at most $2d(v,s)=2d_{\min}$.
The two policies have expected costs at most $\optall$ and
$\opt_D+2p\,d_{\min}=\opt_D(1+2p\beta)$, respectively.  Taking the better
one gives
\[
\frac{\optadapt}{\optpost}
\;\le\; \frac{\min(\rho,\; 1 + 2p\beta)}{(1-p) + p\rho}.
\]

We use three elementary metric constraints.  Shortcutting~$s$ from
$T_{\mathrm{all}}$ gives $\rho\ge1$; inserting~$s$ at a closest vertex of
$T_D$ gives $\rho\le1+2\beta$; and the two edges of $T_{\mathrm{all}}$
incident with~$s$ each have length at least~$d_{\min}$, so
\[
    1\le\rho\le1+2\beta,
    \qquad
    2\beta\le\rho.
\]

For fixed~$\rho$, the numerator $\min(\rho,1+2p\beta)$ is nondecreasing in~$\beta$, and the largest feasible value is $\beta=\rho/2$.
Thus it remains to maximize
\[
    \frac{\min(\rho,1+p\rho)}{(1-p)+p\rho}
    \qquad(\rho\ge1).
\]
If $\rho\le1/(1-p)$, the minimum is~$\rho$ and the expression is increasing in~$\rho$.
If $\rho\ge1/(1-p)$, the minimum is~$1+p\rho$ and the expression is decreasing in~$\rho$.
Therefore the maximum occurs at $\rho=1/(1-p)$, where
\[
\frac{1/(1-p)}{(1-p) + p/(1-p)}
= \frac{1}{(1-p)^2 + p}
= \frac{1}{1 - p + p^2}.
\]

Finally, $1-p+p^2$ is minimized at $p=1/2$.
Hence the largest value of the bound over $p\in[0,1]$ is
\[
    \frac{1}{1-1/2+1/4}=\frac43.
\]
\end{proof}


% ============================================================
% CHAPTER 4: CLAIRVOYANCE GAP — ASYMMETRIC
% ============================================================
\chapter{Clairvoyance Gap for Asymmetric Metrics}\label{ch:asym-clair}

We now turn to asymmetric (directed) metrics and exhibit a family of instances
for which the clairvoyance gap $\optadapt/\optpost$ tends to~$4/3$.

\section{The construction}

We define a family of asymmetric stochastic TSP instances.  Distances are taken to be the directed shortest-path metric induced by the directed edges described below.  The depot is always present.  The client vertices are independent, with probabilities specified below.

For every integer $l\geq 0$, define a directed recursive block $B_l$.  Each block has a distinguished set of permanent vertices, denoted $P_l$.  These permanent vertices will be deterministic clients.  For $l=0$, the block $B_0$ consists of a single permanent vertex.

For $l\geq 1$, the block $B_l$ is obtained from two disjoint copies of $B_{l-1}$, denoted $B_{l-1}^{\mathrm L}$ and $B_{l-1}^{\mathrm R}$, together with a new midpoint vertex $m_l$.  Let
\[
    w_l := 2^{l-1}.
\]
For every permanent vertex $\ell\in P(B_{l-1}^{\mathrm L})$ and every permanent vertex $r'\in P(B_{l-1}^{\mathrm R})$, add a directed edge
$\ell \to r'$
of length $w_l$.  For every $r'\in P(B_{l-1}^{\mathrm R})$, add a directed edge
$r'\to m_l$
of length $w_l$.  For every $\ell\in P(B_{l-1}^{\mathrm L})$, add a directed edge
$m_l\to \ell$
of length $w_l$.  Thus, after contracting the two child blocks, the top level of $B_l$ is the directed triangle
\[
    B_{l-1}^{\mathrm L}
    \longrightarrow
    B_{l-1}^{\mathrm R}
    \longrightarrow
    m_l
    \longrightarrow
    B_{l-1}^{\mathrm L},
\]
where each displayed arc has length $w_l$.

The permanent set of $B_l$ is
$P_l := P(B_{l-1}^{\mathrm L})\cup P(B_{l-1}^{\mathrm R})$.
Thus $|P_l|=2^l$.  All permanent vertices are active with probability~$1$.  Every midpoint vertex is active independently with probability~$1/2$.

The full instance $I_l$ is obtained by adding a depot~$r$.  For every permanent vertex $\ell\in P_l$, add directed edges
$r\to \ell$ and $\ell\to r$,
each of length $w_l=2^{l-1}$.

Let $M_l$ be the set of midpoint vertices in $B_l$.  A midpoint of local depth $k$ has weight $w_k=2^{k-1}$.  There are $2^{l-k}$ midpoint vertices of local depth $k$ inside $B_l$.

\section{Statement}

\begin{theorem}\label{thm:recursive-gap}
For every $l\geq 1$, the instance $I_l$ satisfies
\[
    \frac{\optadapt(I_l)}{\optpost(I_l)}
    =
    \frac{4(l+1)}{3l+4}.
\]
As $l\to\infty$, these ratios approach~$4/3$, so the asymmetric clairvoyance gap is at least~$4/3$.
\end{theorem}

The theorem follows immediately from the following two cost evaluations.  We
state them first and postpone their proofs until after the reduction.

\begin{lemma}[A posteriori cost]\label{lem:post-cost}
For every $l\geq1$,
\[
    \optpost(I_l)=(3l+4)2^{l-2}.
\]
\end{lemma}

\begin{lemma}[Adaptive cost]\label{lem:adapt-cost}
For every $l\geq1$,
\[
    \optadapt(I_l)=(l+1)2^l.
\]
\end{lemma}

\begin{proof}[Proof of Theorem~\ref{thm:recursive-gap}]
By Lemmas~\ref{lem:post-cost} and~\ref{lem:adapt-cost},
\[
    \frac{\optadapt(I_l)}{\optpost(I_l)}
    =
    \frac{(l+1)2^l}{(3l+4)2^{l-2}}
    =
    \frac{4(l+1)}{3l+4}.
\]
The right-hand side tends to~$4/3$ as $l\to\infty$.  Hence the supremum of
the clairvoyance gap over asymmetric instances is at least~$4/3$.
\end{proof}

It remains to prove the two cost lemmas.  Both proofs use the decomposition
tools developed next.

\section{Decomposition tools}

The recursive construction suggests a direct inductive proof.  If a route or
policy served each of the two child blocks in a single uninterrupted visit,
then its cost would split into one service cost for each child and the cost of
the arcs at the current recursive level.  The induction hypothesis would
control the two child costs, leaving only a constant-size top-level
calculation.  A general route or policy, however, may leave a child block,
serve vertices elsewhere, and later re-enter it; visits to the two children
may therefore be interleaved.  To make the induction valid without assuming
contiguous service, we first prove that every additional piece of a child
service can be charged against an arc at the current level.

We call the permanent vertices of a copy of $B_k$ its \emph{ports}.  A
fixed-realization \emph{open service} of $B_k$ starts at a port, serves every
permanent vertex and every active midpoint, and finally terminates at a port.
The initial and terminal ports are free choices.  We allow its expanded walk
to repeat vertices; shortcutting repeated occurrences in the metric shows
that this does not change the fixed-realization optimum.

For the adaptive version we use a slightly relaxed boundary convention that
is convenient for subblocks.  Before its first movement, the policy may call
vertices from an abstract exterior state.  An inactive call leaves it in that
state.  When a call is first active, the policy may choose the port through
which it enters, and only the part of the movement inside $B_k$ is charged.
After this first entry, the usual adaptive rules apply; every still-uncalled
midpoint must be called, and the service finally terminates at a port.  The
terminal movement to that port is charged.  This relaxation models exactly
the internal part of an excursion into a subblock.  Let
\begin{itemize}
    \item $P_k(A)$ be the minimum cost of an open service when the set of
    active midpoints is $A$, and
    \item $Q_k$ be the minimum expected cost of an adaptive open service.
\end{itemize}
Thus $P_0=Q_0=0$.

Whenever a move in the shortest-path metric is made, fix a shortest directed
walk in the generating graph and \emph{expand} the move into that walk.  Every
non-depot arc in the expanded walk belongs to a unique recursive level.
For a child copy $C$, cut the expanded execution whenever it crosses the
boundary of $C$.  The resulting portions that contain service in $C$ are its
\emph{pieces}; each has an entry and an exit port.  Pure transit through $C$
may be assigned to an adjacent piece and only increases its internal cost.

\begin{lemma}[Ports and interrupted services]\label{lem:block-excursions}
Let $C$ be a copy of $B_k$ occurring as a child of a copy of $B_{k+1}$.
Then the following statements hold.
\begin{enumerate}
    \item Any two ports of $C$ can be joined inside $C$ by a directed walk of
    length at most $w_{k+1}$.  For $k\ge1$, this is the identity
    $2w_k=w_{k+1}$; for $k=0$, $C$ has only one port.
    \item Suppose a fixed-realization service of $C$ is split into $N\ge1$
    pieces, each entering and leaving through ports, and let $Z$ be the total
    cost of arcs internal to $C$.  Then
    \[
        Z+w_{k+1}(N-1)\ge P_k(A_C).
        \tag{5.1}\label{eq:post-excursion}
    \]
    \item Suppose an adaptive service of $C$ is split into a random number
    $N\ge1$ of such pieces, possibly according to observations outside $C$.
    If $Z$ is its random internal cost, then
    \[
        \E[Z]+w_{k+1}\E[N-1]\ge Q_k.
        \tag{5.2}\label{eq:adapt-excursion}
    \]
\end{enumerate}
Calls made while the salesperson is outside $C$ that reveal an inactive
vertex of $C$ cause no movement; they may be assigned to the next piece and
do not create an additional piece.
\end{lemma}

\begin{proof}
For the first statement, the case $k=0$ is immediate.  Let $k\ge1$, and let
$x,y$ be ports.  If they lie in different
children of a depth-$j$ block, where $j\le k$ is minimal, then either the arc
from the left child to the right child, of length $w_j$, joins them directly,
or the two-arc walk through that block's midpoint has length $2w_j$.
If they lie in the same child, apply the same argument recursively.  Hence
$d_C(x,y)\le2w_k=w_{k+1}$.

For a fixed realization, concatenate the $N$ pieces in their original order.
Between consecutive pieces, join the exit port of the first to the entry port
of the second by an internal walk of length at most $w_{k+1}$.  The resulting
open service has cost at most $Z+w_{k+1}(N-1)$, which proves
\eqref{eq:post-excursion}.

The same construction works for an adaptive service.  Formally, pre-sample
all activation bits outside $C$ that the original policy observes between two
pieces, and use them only as private random bits when simulating the original
policy inside $C$.  These bits are independent of all activations in $C$.
At the start of a later piece, remain at the preceding exit port while
simulating calls that turn out to be inactive.  When the first active call of
that piece occurs, let $p$ be the port through which its expanded movement
would enter $C$; first move internally to $p$, at cost at most $w_{k+1}$,
and then follow the original internal part of the movement.  Thus only one
port-to-port connector is needed per additional piece.  This gives a valid
randomized open policy for $C$ of expected cost at most the left-hand side of
\eqref{eq:adapt-excursion}.  Randomization cannot improve the optimum of the
finite dynamic program defining $Q_k$: one of the deterministic policies in
its support has no larger expected cost.  This proves
\eqref{eq:adapt-excursion}.
\end{proof}

We next isolate the only top-level calculation needed in the recursion.  At
one recursive node, contract the two children to symbols $L$ and $R$ and
write $M$ for the new midpoint.  The level arcs form
\[
    L\longrightarrow R\longrightarrow M\longrightarrow L,
\]
with every arc having length~$w$.  In the full instance add a symbol $D$ for
the depot and arcs $D\leftrightarrow L$ and $D\leftrightarrow R$, also of
length~$w$.

If the work in $L$ is split into $N_L$ pieces and the work in $R$ into $N_R$
pieces, let $H$ be the number of level arcs used by the expanded execution.
The next lemma deliberately subtracts one arc for every extra piece; these
are precisely the terms paid for by Lemma~\ref{lem:block-excursions}.

\begin{lemma}[Top-level accounting]\label{lem:top-accounting}
For the quotient just described, the following bounds hold.
\begin{enumerate}
    \item For a fixed realization, an open service satisfies
    \[
        H-(N_L-1)-(N_R-1)
        \ge 1+\1_{\{M\text{ active}\}}.
        \tag{5.3}\label{eq:open-top-fixed}
    \]
    A service starting and ending at $D$ satisfies
    \[
        H-(N_L-1)-(N_R-1)
        \ge 3+\1_{\{M\text{ active}\}}.
        \tag{5.4}\label{eq:closed-top-fixed}
    \]
    \item For an adaptive service, with $M$ active independently with
    probability $1/2$,
    \[
        \E\bigl[H-(N_L-1)-(N_R-1)\bigr]\ge2
        \tag{5.5}\label{eq:open-top-adapt}
    \]
    in the open case, and
    \[
        \E\bigl[H-(N_L-1)-(N_R-1)\bigr]\ge4
        \tag{5.6}\label{eq:closed-top-adapt}
    \]
    in the case starting and ending at $D$.
\end{enumerate}
\end{lemma}

\begin{proof}
Associate with every piece of $L$ except its first the last level arc by which
the execution enters $L$ before that piece; do the same for $R$.  Calls made
without movement to inactive vertices are assigned to an adjacent piece, so
every extra piece has such an entering arc.  The associated arcs are
distinct.  Call all remaining level arcs \emph{uncharged}.  Their number is
exactly
\[
    \widehat H:=H-(N_L-1)-(N_R-1).
\]

In an open fixed-realization service with $M$ inactive, the first service of
whichever child is not the starting child requires an uncharged entering arc;
hence $\widehat H\ge1$.  If $M$ is active, the first entry into that other
child and the entry into $M$ are two distinct uncharged arcs, so
$\widehat H\ge2$.  This proves \eqref{eq:open-top-fixed}.  In a service that
starts and ends at $D$, the first entries used to serve $L$ and $R$, together
with the final return to $D$, are three distinct uncharged arcs.  When $M$ is
active, its entry is a fourth.  This proves
\eqref{eq:closed-top-fixed}.

For the adaptive bounds, condition on all activation bits except that of $M$
and couple the two resulting executions.  Write $\widehat H_0$ and
$\widehat H_1$ for the uncharged counts when $M$ is respectively inactive and
active.  The fixed-realization bounds give
\[
    \widehat H_0\ge1,
    \qquad
    \widehat H_1\ge2
\]
in the open case.  If $\widehat H_0\ge2$, their sum is already at least~$4$.
It remains to consider $\widehat H_0=1$.  Equality forces the inactive
execution to use its sole uncharged arc to go from a first service of $L$ to a
first service of $R$; the reverse direction needs two cycle arcs.  Thus the
uncharged skeleton is $L\to R$.  If $M$ is called before this transition, the
active execution must enter $M$ and must still serve $R$ and terminate at a
port, using at least three uncharged arcs.  If $M$ is called after the
transition, the two executions have already used $L\to R$, and the active
execution additionally needs $R\to M\to L$, again giving
$\widehat H_1\ge3$.  The same conclusion holds under the relaxed boundary
convention: calling $M$ before entering either child costs one arc when active,
and completing $M\to L\to R$ costs two more.  Hence in every case
\[
    \widehat H_0+\widehat H_1\ge4.
\]
Since the two values of $M$ are equiprobable, this proves
\eqref{eq:open-top-adapt}.

In the closed case, \eqref{eq:closed-top-fixed} gives
$\widehat H_0\ge3$ and $\widehat H_1\ge4$.  If
$\widehat H_0\ge4$, their sum is at least~$8$.  If
$\widehat H_0=3$, equality in the first bound forces the inactive uncharged
skeleton to be
\[
    D\to L\to R\to D;
\]
all other orders require at least four arcs.  If $M$ is called before the two
child services are complete, the active execution must enter $M$ and still
complete the missing first entries and the return to $D$.  More explicitly,
calling $M$ at $D$ before departure, at $L$, at $R$ before the final return,
or at $D$ after the return requires at least $5,6,5,7$ uncharged arcs,
respectively, in the active execution.  Thus $\widehat H_1\ge5$.  Therefore
$\widehat H_0+\widehat H_1\ge8$, which proves
\eqref{eq:closed-top-adapt}.
\end{proof}

\section{Proof of the a posteriori cost lemma}

We can now justify the recursive formula without assuming that the two child
blocks are served contiguously.

\begin{lemma}\label{lem:post-rec}
For every $k\ge1$ and every realization $A\subseteq M_k$,
\[
    P_k(A)
    =P_{k-1}(A_{\mathrm L})+P_{k-1}(A_{\mathrm R})
     +w_k\bigl(1+\1_{\{m_k\in A\}}\bigr).
    \tag{5.7}\label{eq:post-rec}
\]
\end{lemma}

\begin{proof}
Expand an arbitrary open service.  Let $Z_{\mathrm L},Z_{\mathrm R}$ be
the costs of arcs internal to the two children, let
$N_{\mathrm L},N_{\mathrm R}$ be their numbers of pieces, and let $H$ be the
number of top-level arcs.  Lemmas~\ref{lem:block-excursions}
and~\ref{lem:top-accounting} give
\begin{align*}
    \cost
    &=Z_{\mathrm L}+Z_{\mathrm R}+w_kH\\
    &\ge P_{k-1}(A_{\mathrm L})+P_{k-1}(A_{\mathrm R})
       +w_k\bigl(H-(N_{\mathrm L}-1)-(N_{\mathrm R}-1)\bigr)\\
    &\ge P_{k-1}(A_{\mathrm L})+P_{k-1}(A_{\mathrm R})
       +w_k\bigl(1+\1_{\{m_k\in A\}}\bigr).
\end{align*}
For the reverse inequality, if $m_k$ is inactive, optimally serve the left
child and then the right child.  If $m_k$ is active, optimally serve the right
child, traverse $R\to m_k\to L$, and then serve the left child.  Because the
ports at both ends are free, these child services concatenate without any
additional internal cost.  This attains the displayed lower bound.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:post-cost}]
Let $p_k:=\E[P_k(A)]$.  Since $m_k$ is active with probability $1/2$,
\[
    p_k=2p_{k-1}+\frac32w_k,
    \qquad p_0=0.
\]
As there are $2^{k-j}$ depth-$j$ midpoints and every such midpoint has weight
$w_j=2^{j-1}$,
\[
    p_k
    =\frac32\sum_{j=1}^k2^{k-j}w_j
    =\frac32\sum_{j=1}^k2^{k-1}
    =3k2^{k-2}.
    \tag{5.8}\label{eq:expected-open-post}
\]

It remains to add the depot.  Expanded client-to-client shortest paths may be
chosen not to use the depot: if such a path contains
$x\to r\to y$ with $x,y\in P_l$, replace these two depot arcs, of total
length $2w_l$, by an internal port-to-port path of length at most $2w_l$.
Thus a tour uses one depot arc when it first enters $B_l$ and one when it
finally leaves.  Removing those two arcs leaves an open service, and conversely
an open service can be closed using two depot arcs.  Hence, for every
realization,
\[
    \opt(A)=2w_l+P_l(A).
\]
Taking expectations and using \eqref{eq:expected-open-post},
\[
    \optpost(I_l)
    =2^l+3l2^{l-2}
    =(3l+4)2^{l-2}.
    \tag{5.9}\label{eq:post-value-recursive}
\]
\end{proof}

\section{Proof of the adaptive cost lemma}

\begin{lemma}\label{lem:adapt-rec}
For every $k\ge1$,
\[
    Q_k=2Q_{k-1}+2w_k.
    \tag{5.10}\label{eq:adapt-rec}
\]
Consequently $Q_k=k2^k$.
\end{lemma}

\begin{proof}
Consider an arbitrary adaptive open service and use the same notation as in
the proof of Lemma~\ref{lem:post-rec}.  By
\eqref{eq:adapt-excursion} and \eqref{eq:open-top-adapt},
\begin{align*}
    \E[\cost]
    &=\E[Z_{\mathrm L}+Z_{\mathrm R}+w_kH]\\
    &\ge 2Q_{k-1}
      +w_k\E\bigl[H-(N_{\mathrm L}-1)-(N_{\mathrm R}-1)\bigr]\\
    &\ge2Q_{k-1}+2w_k.
\end{align*}

For the matching policy, serve the left child, then the right child, and then
call $m_k$.  If $m_k$ is inactive, terminate at the current right-child port.
If it is active, traverse $R\to m_k\to L$ and terminate at a left-child port.
The top-level expected cost is
$w_k+\frac12(2w_k)=2w_k$, so equality holds in
\eqref{eq:adapt-rec}.  Induction, using $Q_0=0$ and $w_k=2^{k-1}$, gives
$Q_k=k2^k$.
\end{proof}

\begin{proof}[Proof of Lemma~\ref{lem:adapt-cost}]
For the full instance, the same decomposition applies, except that the
top-level quotient starts and ends at the depot.  Lemma~\ref{lem:top-accounting}
therefore gives the lower bound
\[
    \optadapt(I_l)\ge2Q_{l-1}+4w_l.
\]
It is attained by the recursive policy that serves the left child, then the
right child, and then calls $m_l$: if $m_l$ is inactive it returns directly
from the right child to the depot, while if $m_l$ is active it follows
$R\to m_l\to L\to r$.  Hence
\begin{align*}
    \optadapt(I_l)
    &=2Q_{l-1}+4w_l\\
    &=2(l-1)2^{l-1}+4\cdot2^{l-1}\\
    &=(l+1)2^l.
    \tag{5.11}\label{eq:adapt-value-recursive}
\end{align*}
\end{proof}


% ============================================================
% CHAPTER 5: APPROXIMATE A PRIORI TSP
% ============================================================
\chapter{Approximate A Priori TSP via Dynamic Programming}\label{ch:dp}

We give a deterministic $(1+\varepsilon)$-approximation algorithm for
a~priori TSP on directed metrics.  For every fixed $\varepsilon>0$, its
running time is $c_\varepsilon^n\poly(n)$.

The algorithm is based on the following view of an a~priori order.  Once an
order $\pi=(r,v_1,\dots,v_n,r)$ is fixed, an edge from $v_i$ to $v_j$ is paid
exactly when $v_i$ and $v_j$ are active and all clients between them in the
order are inactive.  Thus the expected cost is a sum over ordered pairs.  The
coefficient of a pair contains the product
\[
    Q(i,j)=\prod_{i<k<j}(1-p_k),
\]
the probability that the interval between the two endpoints is completely
inactive.

Pairs with very small $Q(i,j)$ correspond to rare long shortcuts.  The first
part of the chapter proves that, up to a factor $1+O(\varepsilon)$, such pairs
may be discarded.  The proof does not try to charge each discarded pair
separately.  Instead, it slightly increases all activation probabilities.  This
adds enough active clients to split long jumps, while the directed triangle
inequality ensures that the shortcut cost can only increase.  This gives a
uniform bound comparing the full cost with the truncated pairwise cost.

It remains to minimize this truncated objective.  Directly remembering the
entire prefix of a partial order would be too expensive, so we group clients
into blocks: high-probability clients are kept as singletons, while
low-probability clients are grouped into small-mass blocks.  For each block we
precompute the best internal order.  Between two blocks, the truncated objective
depends only on the complete blocks between them.  Consequently, when the
dynamic program builds the block order from left to right, blocks that are
separated from the current end by inactivity product at most~$\eta$ can never
interact with future blocks and can be forgotten.

The rest of the chapter follows this outline: first the pairwise objective,
then the truncation lemma, then the block objective and its comparison with the
original order.  After that we give the dynamic program and choose the
parameters that yield the $(1+\varepsilon)$ guarantee.

\section{The pairwise objective}

Let
\[
    \pi=(r,v_1,\dots,v_n,r)
\]
be an a~priori order.  Put $v_0=v_{n+1}=r$ and $p_0=p_{n+1}=1$.  For brevity,
write
\[
    p_i:=p_{v_i},
    \qquad
    d_{ij}:=d(v_i,v_j).
\]
For $0\le i<j\le n+1$, define
\[
    Q(i,j):=\prod_{i<k<j}(1-p_k).
\]
Thus $Q(i,j)$ is the probability that every client strictly between $v_i$ and
$v_j$ is inactive.  The expected cost of $\pi$ is
\begin{equation}\label{eq:apriori-pairwise}
    \cost(\pi)
    =
    \sum_{0\le i<j\le n+1}p_i p_j Q(i,j)d_{ij}.
\end{equation}
Indeed, the term indexed by $(i,j)$ is paid exactly when $v_i$ and $v_j$ are
consecutive active vertices in the shortcut of~$\pi$.

Fix $0<\eta<1$.  We retain a pair $(i,j)$ when
$Q(i,j)>\eta$, and define
\begin{equation}\label{eq:local-pair-cost}
    \cost^{\loc}_{\eta}(\pi)
    :=
    \sum_{\substack{0\le i<j\le n+1\\Q(i,j)>\eta}}
    p_i p_j Q(i,j)d_{ij}.
\end{equation}

\section{\texorpdfstring{Truncating pairs with small \(Q\)}{Truncating pairs with small Q}}

We avoid choosing a splitting vertex separately for every discarded pair.
Instead, we slightly increase all activation probabilities.  The additional
active clients split all long jumps simultaneously, and the directed triangle
inequality shows that this can only increase the shortcut cost.

The parameter $\alpha$ measures the strength of this additional activation
round.  Write $q_v:=1-p_v$.  Besides the original active set, independently
activate $v$ with probability $1-q_v^\alpha$.  Then $v$ is absent from both
rounds with probability
\[
    q_v q_v^\alpha=q_v^{1+\alpha}.
\]
Consequently, a gap whose original inactivity probability is $Q(i,j)$ remains
empty after the additional round with probability
\[
    Q(i,j)^{1+\alpha}=Q(i,j)\,Q(i,j)^\alpha.
\]
Thus every discarded pair with $Q(i,j)\le\eta$ gains the uniform damping factor
$Q(i,j)^\alpha\le\eta^\alpha$.  The price is that each endpoint activation
probability grows by at most a factor $1+\alpha$.  The proof below formalizes
this tradeoff between the endpoint factor $(1+\alpha)^2$ and the damping factor
$\eta^\alpha$.

For an activation vector $s=(s_v)_{v\in V}$, let $\cost_s(\pi)$ denote the
expected cost of~$\pi$ under independent activation probabilities~$s_v$.
Thus $\cost_p(\pi)=\cost(\pi)$ for the original vector~$p$.

\begin{lemma}[Truncation by increasing activation probabilities]
\label{lem:long-gap-truncation}
Let $\alpha>0$, put $\kappa:=(1+\alpha)^2$, and suppose that
$0<\eta<1$ and $\kappa\eta^\alpha<1$.  Then every a~priori order~$\pi$
satisfies
\begin{equation}\label{eq:long-gap-truncation}
    \cost(\pi)
    \le
    \frac{\kappa}{1-\kappa\eta^\alpha}\,
    \cost^{\loc}_{\eta}(\pi).
\end{equation}
\end{lemma}

\begin{proof}
Define
\[
    \widetilde p_v:=1-(1-p_v)^{1+\alpha}.
\]
To couple the two activation vectors, first sample the original active set~$A$.
Independently, form a set~$B$ by including each client~$v$ with probability
$1-(1-p_v)^\alpha$.  Then $\widetilde A:=A\cup B$ has independent activation
probabilities~$\widetilde p_v$.

For every outcome, $A\subseteq\widetilde A$, and $\pi\vert_A$ is obtained from
$\pi\vert_{\widetilde A}$ by shortcutting vertices.  Hence the directed
triangle inequality gives
\[
    \cost(\pi\vert_A)\le \cost(\pi\vert_{\widetilde A}).
\]
Taking expectations,
\begin{equation}\label{eq:activation-monotonicity}
    \cost(\pi)\le \cost_{\widetilde p}(\pi).
\end{equation}

Convexity of $x\mapsto x^{1+\alpha}$ gives
\[
    \widetilde p_v
    =1-(1-p_v)^{1+\alpha}
    \le(1+\alpha)p_v.
\]
Moreover,
\[
    \widetilde Q(i,j)
    =\prod_{i<k<j}(1-\widetilde p_k)
    =Q(i,j)^{1+\alpha}.
\]
Thus the coefficient of $d_{ij}$ under~$\widetilde p$ is at most
\begin{align*}
    \widetilde p_i\widetilde p_j\widetilde Q(i,j)
    &\le
    \kappa p_ip_jQ(i,j)^{1+\alpha}\\
    &=\kappa Q(i,j)^\alpha\bigl(p_ip_jQ(i,j)\bigr).
\end{align*}

Set
\[
    T:=\cost(\pi),
    \qquad
    C:=\cost^{\loc}_{\eta}(\pi).
\]
On retained pairs, $Q(i,j)^\alpha\le1$; on discarded pairs,
$Q(i,j)^\alpha\le\eta^\alpha$.  Therefore
\[
    T
    \le \cost_{\widetilde p}(\pi)
    \le \kappa\bigl(C+\eta^\alpha(T-C)\bigr)
    \le \kappa C+\kappa\eta^\alpha T.
\]
Rearranging proves~\eqref{eq:long-gap-truncation}.
\end{proof}

\section{The block objective}

The dynamic program will order blocks rather than individual clients.  We
therefore define an objective on admissible ordered partitions.  For endpoints
in different blocks, this objective retains the exact inactivity factors of all
complete intervening blocks but replaces the factors contributed by vertices
inside the two endpoint blocks by~$1$.  This gives an upper bound on the
retained pair contribution and allows the contribution of a new block to be
evaluated from a bounded suffix of the preceding blocks.

Fix $0<\delta<1/4$.  A client is \emph{heavy} if
$p_v>\delta/2$ and \emph{light} otherwise.  For a set of clients~$B$, write
\[
    \mu(B):=\sum_{v\in B}p_v,
    \qquad
    q(B):=\prod_{v\in B}(1-p_v).
\]
For the depot block $\{r\}$, set $q(\{r\})=1$.

\begin{definition}[Admissible blocks and partitions]\label{def:admissible-partition}
A block is \emph{admissible} if it is either
\begin{enumerate}[label=(\alph*)]
    \item a singleton containing one heavy client, or
    \item a nonempty set of light clients with total mass at most~$\delta$.
\end{enumerate}
An ordered partition $(B_1,\dots,B_m)$ is admissible if all its blocks are
admissible and, whenever $B_{t-1}$ and $B_t$ are both light blocks,
\[
    \mu(B_t)\ge\delta/2.
\]
Thus, in every consecutive run of light blocks, only the first block may have
mass below~$\delta/2$.
\end{definition}

For a light block~$B$, define
\begin{equation}\label{eq:internal-block-cost}
    I(B)
    :=
    \min_{\sigma}
    \sum_{x<_\sigma y}p_xp_y d(x,y),
\end{equation}
where the minimum is over all orders~$\sigma$ of~$B$.  For a heavy singleton
and for the depot block, set $I=0$.  The values $I(B)$, together with a
minimizing order, can be computed by the recurrence
\begin{equation}\label{eq:internal-block-recurrence}
    I(B)
    =
    \min_{y\in B}
    \left(
        I(B\setminus\{y\})
        +\sum_{x\in B\setminus\{y\}}p_xp_y d(x,y)
    \right),
    \qquad I(\emptyset)=0.
\end{equation}
Indeed, the second term is exactly the contribution of the pairs whose last
vertex is~$y$.

Given an admissible ordered partition
$\mathcal B=(B_1,\dots,B_m)$, order each light block according to a minimizer in
\eqref{eq:internal-block-cost}, and concatenate the block orders.  Denote the
resulting a~priori order by~$\pi(\mathcal B)$.  Put
$B_0=B_{m+1}=\{r\}$ and $p_r=1$, and define
\[
    C(B_i,B_j)
    :=
    \sum_{x\in B_i}\sum_{y\in B_j}p_xp_y d(x,y).
\]
Define the \emph{block objective} by
\begin{equation}\label{eq:block-objective}
    F_\eta(\mathcal B)
    :=
    \sum_{t=1}^m I(B_t)
    +
    \sum_{\substack{0\le i<j\le m+1\\
                     \prod_{i<t<j}q(B_t)>\eta}}
    \left(\prod_{i<t<j}q(B_t)\right)C(B_i,B_j).
\end{equation}
The first sum accounts for pairs whose endpoints lie in the same block.  The
second sum accounts for pairs in different blocks whenever the inactivity
product of the complete intervening blocks exceeds~$\eta$.

\begin{lemma}[Comparison with the exact objective]\label{lem:block-comparison}
The block objective satisfies the following two inequalities.
\begin{enumerate}[label=(\roman*)]
    \item For every admissible ordered partition~$\mathcal B$,
    \begin{equation}\label{eq:block-objective-dominates-local}
        \cost^{\loc}_{\eta}(\pi(\mathcal B))
        \le F_\eta(\mathcal B).
    \end{equation}

    \item For every a~priori order~$\pi$, there is an admissible ordered
    partition $\mathcal B_\pi$ into consecutive sets of~$\pi$ such that
    \begin{equation}\label{eq:block-objective-from-order}
        F_\eta(\mathcal B_\pi)
        \le(1-\delta)^{-2}\cost(\pi).
    \end{equation}
\end{enumerate}
\end{lemma}

\begin{proof}
For~(i), first consider endpoints $x\in B_i$ and $y\in B_j$ with $i<j$.
In the order $\pi(\mathcal B)$, their exact coefficient is
\[
    p_xp_y\,
    q(B_i^{>x})\,
    \left(\prod_{i<t<j}q(B_t)\right)\,
    q(B_j^{<y}),
\]
where $B_i^{>x}$ denotes the vertices of $B_i$ after~$x$, and
$B_j^{<y}$ the vertices of $B_j$ before~$y$.  If $(x,y)$ is retained in
$\cost^{\loc}_{\eta}$, then its exact inactivity probability exceeds
$\eta$.  Therefore the product over the complete intervening blocks also
exceeds $\eta$, so the corresponding block pair occurs in
\eqref{eq:block-objective}.  Dropping the two endpoint factors can only
increase its coefficient.  If both endpoints lie in the same light block~$B$,
then their exact coefficient is at most $p_xp_y$, and the sum of these upper
bounds in the chosen internal order is exactly~$I(B)$.  Summing over all
retained pairs proves~\eqref{eq:block-objective-dominates-local}.

For~(ii), keep every heavy client as a singleton.  In each maximal consecutive
run of light clients in~$\pi$, scan from right to left.  Close a block as soon
as its mass reaches~$\delta/2$, and continue with the remaining prefix of the
run.  Every closed block has mass in $[\delta/2,\delta]$, because every light
client has probability at most~$\delta/2$.  The only possible block of mass
below~$\delta/2$ is the leftmost block of the run.  The resulting partition
$\mathcal B_\pi$ is therefore admissible.

Consider a cross-block term of $F_\eta(\mathcal B_\pi)$ with endpoints
$x\in B_i$ and $y\in B_j$, where $i<j$.  Relative to the exact coefficient of
this pair in~$\pi$, the corresponding term of $F_\eta(\mathcal B_\pi)$ omits
only the factors corresponding to the vertices after~$x$ in~$B_i$ and
before~$y$ in~$B_j$.  For a heavy singleton or
a depot block, the corresponding factor is~$1$.  For a light block~$B$,
\[
    q(B')\ge1-\sum_{v\in B'}p_v\ge1-\delta
    \qquad\text{for every }B'\subseteq B.
\]
Thus every cross-block coefficient in $F_\eta(\mathcal B_\pi)$ is at most
$(1-\delta)^{-2}$ times the corresponding exact coefficient in~$\pi$.

Finally, let $B$ be a light block, and retain the order induced by~$\pi$ on~$B$.
If $T_B(\pi)$ denotes the exact contribution of pairs with both endpoints
in~$B$, then
\[
    I(B)
    \le
    \sum_{\substack{x,y\in B\\x<_\pi y}}p_xp_y d(x,y)
    \le
    \frac{1}{1-\delta}T_B(\pi).
\]
The second inequality holds because every inactivity factor between two
vertices of~$B$ is at least $q(B)\ge1-\delta$.  Summing the cross-block and
internal comparisons proves~\eqref{eq:block-objective-from-order}.
\end{proof}

\section{The dynamic program}

We now minimize $F_\eta$ exactly over all admissible ordered partitions.
Consider a partial partition
\[
    (B_0,B_1,\dots,B_t),
    \qquad B_0=\{r\}.
\]
If
\[
    \prod_{s=1}^t q(B_s)>\eta,
\]
its \emph{relevant suffix} is the entire prefix
$(B_0,B_1,\dots,B_t)$.  Otherwise, let $a$ be the largest index such that
\[
    \prod_{s=a}^t q(B_s)\le\eta.
\]
Equivalently,
\[
    \prod_{s=a}^t q(B_s)\le\eta
    \quad\text{and}\quad
    \prod_{s=a+1}^t q(B_s)>\eta.
\]
The relevant suffix is then $(B_a,B_{a+1},\dots,B_t)$.

A state consists of the set $U\subseteq V$ of already placed clients and the
relevant suffix~$\Omega$ of their partial partition.  Its value is the minimum,
over all represented prefixes, of the internal terms of the placed blocks plus
all cross-block terms whose right endpoint has already been placed.  The
initial state is $(\emptyset,(B_0))$ with value~$0$.

A feasible next block~$B$ is either a singleton containing an unplaced heavy
client or a nonempty set of unplaced light clients with
$\mu(B)\le\delta$.  If the last block of the current prefix and~$B$ are both
light, we additionally require $\mu(B)\ge\delta/2$.  Suppose
$\Omega=(B_a,\dots,B_t)$.  Appending~$B$ adds
\begin{equation}\label{eq:dp-increment}
    I(B)
    +
    \sum_{s=a}^t
    \left(\prod_{s<u\le t}q(B_u)\right)C(B_s,B).
\end{equation}
Every block in the relevant suffix occurs in this sum, because the complete
blocks between it and~$B$ have inactivity product greater than~$\eta$.
After the transition, the suffix is trimmed according to the same rule.  Once
every client has been placed, the terminal depot block~$\{r\}$ is appended
using the cross-block part of~\eqref{eq:dp-increment} and internal cost zero.

\begin{lemma}[Correctness and running time of the DP]\label{lem:local-dp}
For fixed $\delta$ and~$\eta$, the dynamic program minimizes
$F_\eta$ over all admissible ordered partitions and runs in time
$c_{\delta,\eta}^n\poly(n)$.
\end{lemma}

\begin{proof}
When a block~$B$ is appended, the only new terms of~$F_\eta$ are its internal
term $I(B)$ and the cross-block terms whose right endpoint is~$B$.  These are
exactly the terms in~\eqref{eq:dp-increment}.

If a previous block is not in the relevant suffix, then the complete blocks
between it and the current end have inactivity product at most~$\eta$.
Appending more blocks only multiplies this product by numbers in~$[0,1]$.
Hence a forgotten block can never occur in a future term of~$F_\eta$.  The
state therefore contains precisely the information needed to evaluate every
continuation, and the recurrence is an exact shortest-path computation over all
admissible partial partitions.

It remains to bound the number of states.  We use the elementary inequality
\[
    \prod_{v\in S}(1-p_v)
    \le \exp\left(-\sum_{v\in S}p_v\right).
\]
If the whole prefix is retained, then its inactivity product exceeds
$\eta$, so its total activation mass is less than
$\log(1/\eta)$.  Otherwise, deleting the first retained block leaves an
inactivity product greater than~$\eta$, so the remaining blocks have total
activation mass less than $\log(1/\eta)$; the first retained block has mass at
most~$1$.  Thus the total activation mass of a relevant suffix is less than
\[
    M_\eta:=\log(1/\eta)+1.
\]

Every retained heavy singleton has mass greater than~$\delta/2$, and every
retained light block except the first one in a light run has mass at least
$\delta/2$.  Let $\mathcal G$ be the collection of these blocks.  Since their
total mass is less than $M_\eta$,
\[
    |\mathcal G|<2\frac{M_\eta}{\delta}.
\]
Every remaining retained block is the first light block of a light run.  The
number of light runs is at most one more than the number of retained heavy
blocks, and hence at most $|\mathcal G|+1$.  Therefore the number of retained
client blocks is at most
\[
    2|\mathcal G|+1
    <4\frac{M_\eta}{\delta}+1.
\]
Including the depot block, the relevant suffix has at most
\[
    K:=\left\lceil4\frac{M_\eta}{\delta}\right\rceil+2
    =O\!\left(\frac{\log(1/\eta)+1}{\delta}\right)
\]
blocks.  A state can be encoded by assigning each client one of the labels
``unplaced'', ``placed and forgotten'', or ``position $1,\dots,K$ in the
relevant suffix''.  Thus there are at most $(K+2)^n\poly(n)$ states.  A
transition may enumerate a subset of the unplaced light clients, contributing
at most another factor~$2^n$, and all numerical terms can be evaluated in
polynomial time.  This gives a running time
$c_{\delta,\eta}^n\poly(n)$.  The recurrence
\eqref{eq:internal-block-recurrence} for the values $I(B)$ takes
$O(n2^n)$ time and is absorbed in the same bound.
\end{proof}

\section{Approximation guarantee}

\begin{theorem}\label{thm:directed-apriori-ptas}
For every fixed $0<\varepsilon\le1$, a~priori TSP on directed
metrics has a deterministic algorithm running in time
$c_\varepsilon^n\poly(n)$ and returning an a~priori order~$\hat\pi$ such that
\[
    \cost(\hat\pi)\le(1+\varepsilon)\optprior.
\]
\end{theorem}

\begin{proof}
Set
\[
    \delta:=\frac{\varepsilon}{16},
    \qquad
    \alpha:=\frac{\varepsilon}{32},
    \qquad
    \eta:=\left(\frac{\varepsilon}{64}\right)^{32/\varepsilon}.
\]
Then
\[
    \eta^\alpha=\frac{\varepsilon}{64}.
\]

Let $\pi^*$ be an optimal a~priori order.  By
Lemma~\ref{lem:block-comparison}(ii), there is an admissible ordered partition
$\mathcal B^*$ into consecutive sets of~$\pi^*$ such that
\[
    F_\eta(\mathcal B^*)
    \le(1-\delta)^{-2}\cost(\pi^*)
    =(1-\delta)^{-2}\optprior.
\]
Let $\widehat{\mathcal B}$ be a partition minimizing $F_\eta$, as computed by
Lemma~\ref{lem:local-dp}, and put
$\hat\pi:=\pi(\widehat{\mathcal B})$.  Lemma~\ref{lem:block-comparison}(i)
gives
\[
    \cost^{\loc}_{\eta}(\hat\pi)
    \le F_\eta(\widehat{\mathcal B})
    \le F_\eta(\mathcal B^*)
    \le(1-\delta)^{-2}\optprior.
\]
Applying Lemma~\ref{lem:long-gap-truncation},
\begin{equation}\label{eq:final-approximation-chain}
    \cost(\hat\pi)
    \le
    \frac{(1+\alpha)^2}
         {1-(1+\alpha)^2\eta^\alpha}
    (1-\delta)^{-2}\optprior.
\end{equation}

Put $\kappa:=(1+\alpha)^2$.  Since $0<\varepsilon\le1$,
\[
    \kappa
    =1+\frac{\varepsilon}{16}+\frac{\varepsilon^2}{1024}
    \le1+\frac{\varepsilon}{8}
    \le\frac98.
\]
Therefore
\[
    \kappa\eta^\alpha
    =\kappa\frac{\varepsilon}{64}
    \le\frac{9\varepsilon}{512}
    <\frac{\varepsilon}{32}.
\]
Using $(1-x)^{-1}\le1+2x$ for $0\le x\le1/2$, we get
\begin{align*}
    \frac{\kappa}{1-\kappa\eta^\alpha}
    &\le
    \left(1+\frac{\varepsilon}{8}\right)
    \left(1+\frac{\varepsilon}{16}\right)\\
    &\le 1+\frac{25\varepsilon}{128}
    <1+\frac{\varepsilon}{4}.
\end{align*}
Also,
\[
    (1-\delta)^{-2}
    \le(1-2\delta)^{-1}
    =\frac{1}{1-\varepsilon/8}
    \le1+\frac{\varepsilon}{4}.
\]
Substituting these bounds into~\eqref{eq:final-approximation-chain} yields
\[
    \cost(\hat\pi)
    \le
    \left(1+\frac{\varepsilon}{4}\right)^2\optprior
    \le(1+\varepsilon)\optprior.
\]
Finally,
\[
    \log(1/\eta)
    =\frac{32}{\varepsilon}\log\frac{64}{\varepsilon},
\]
so $\delta$ and~$\eta$ depend only on~$\varepsilon$.  The running time remains
$c_\varepsilon^n\poly(n)$.
\end{proof}

\chapter{Conclusion and Future Work}\label{ch:conclusion}

\section{Summary of results}

We summarize the main results of this thesis in the following table.

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Result} & \textbf{Metric} & \textbf{Reference} \\
\midrule
Adaptivity gap $\ge 5/4$ & Symmetric & Theorem~\ref{thm:cycle-subdivision} \\
Clairvoyance gap $\ge 11/10$ & Symmetric & Theorem~\ref{thm:sym-clair} \\
One stochastic vertex: gap $\le 1/(1-p+p^2)$ & Symmetric & Proposition~\ref{prop:general-p} \\
Clairvoyance gap $\ge 4/3$ & Asymmetric & Theorem~\ref{thm:recursive-gap} \\
$(1{+}\varepsilon)$-approx.\ for a~priori TSP & Asymmetric & Theorem~\ref{thm:directed-apriori-ptas} \\
\bottomrule
\end{tabular}
\end{center}

\section{Directions for future research}

The results of this thesis suggest several possible directions for subsequent
work.

\begin{enumerate}
    \item \textbf{Symmetric clairvoyance gap.}
    What is the correct supremum of the clairvoyance gap for symmetric stochastic TSP?
    The three-chain construction gives a lower bound of~$11/10$, while Proposition~\ref{prop:general-p} gives an upper bound of~$4/3$ for the special case of one stochastic vertex.

    \item \textbf{Symmetric adaptivity gap.}
    Our lower bound of $5/4$ is the best known.  Is the adaptivity gap for symmetric metrics strictly larger?  What is its exact supremum?

    \item \textbf{Asymmetric clairvoyance gap.}
    Is the clairvoyance gap for asymmetric metrics $\Theta(1)$?  Our recursive construction gives a lower bound tending to~$4/3$.  Can this be matched by an $O(1)$ upper bound, or does the gap grow with~$n$?

    \item \textbf{Efficient adaptive policies.}
    Can one design polynomial-time adaptive policies whose expected cost is within a constant factor of~$\optadapt$ for asymmetric metrics?
    A promising direction is the backbone circulation approach: sample a backbone, construct a low-cost fractional circulation, round it to an integral assignment, and convert the result into an adaptive policy.

    \item \textbf{Faster a~priori approximation algorithms.}
    Our $(1{+}\varepsilon)$-approximation for directed a~priori TSP runs in
    time $c_\varepsilon^n\poly(n)$ and is therefore exponential in~$n$.
    A polynomial-time $(1{+}\varepsilon)$-approximation scheme would also
    apply when every vertex has activation probability~$1$, and would thus
    give a PTAS for the classical directed TSP.  Since that problem is
    APX-hard, this is impossible unless $\mathrm{P}=\mathrm{NP}$~\cite{Karpinski2015}.
    A more realistic direction is
    to seek a polynomial-time constant-factor approximation for directed
    a~priori TSP, or more generally to improve the tradeoff between running
    time and approximation ratio.
\end{enumerate}


% ============================================================
% ACKNOWLEDGEMENTS
% ============================================================
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank Dr.\ Sharat Ibrahimpur for his guidance throughout this
project, Manuel Christalla for the continuous discussions and valuable
suggestions, and Prof.\ Dr.\ Vera Traub for advising this thesis.

I would also like to give special thanks to my parents and my girlfriend for
their constant care, understanding, and support in my life.


% ============================================================
% APPENDIX: EXPLORATORY COMPUTATIONAL SEARCH
% ============================================================
\appendix
\chapter{Exploratory Computational Search}\label{app:evidence}

This appendix records computational searches used as exploratory guidance for the symmetric clairvoyance gap.
The search procedures were heuristic and limited in scope, so the table below should be read only as a record of tested instances, not as proof of tightness.

Simulated-annealing search over continuous symmetric distance matrices and integer-weight sparse graphs produced the following best ratios:

\begin{center}
\begin{tabular}{cllll}
\toprule
$n$ & Best ratio & Exact fraction & Source & Notes \\
\midrule
4 & $1.0000$ & $1$ & exact & no gap possible \\
6 & $1.0667$ & $16/15$ & SA (int.) & verified by DP \\
8 & $1.0909$ & $12/11$ & SA (int.) & verified by DP \\
10 & $1.1000$ & $11/10$ & 3-chain $k=3$ & verified by DP \\
13 & $1.1000$ & $11/10$ & 3-chain $k=4$ & verified by DP \\
16 & $1.1000$ & $11/10$ & 3-chain $k=5$ & verified by DP \\
19 & $1.1000$ & $11/10$ & 3-chain $k=6$ & verified by DP \\
22 & $1.1000$ & $11/10$ & 3-chain $k=7$ & verified by DP \\
25 & $1.1000$ & $11/10$ & 3-chain $k=8$ & verified by DP \\
\bottomrule
\end{tabular}
\end{center}

The search did not find a ratio above $11/10$ in these runs.
In the listed instances attaining $11/10$, the construction has a single stochastic vertex with activation probability $1/2$.


% ============================================================
% APPENDIX: DECLARATION OF ORIGINALITY
% ============================================================
\chapter{Declaration of Originality}\label{app:declaration}

I hereby declare that I authored this thesis independently.  Suggestions from
the supervisor regarding language and content are excepted.  All sources and
aids used in preparing this thesis, including any authorized generative
artificial intelligence technologies, are documented truthfully and fully in
consultation with the supervisor.  I have adhered to the applicable citation
guidelines, and all passages taken verbatim or paraphrased from published or
unpublished works of others are indicated as such.

\bigskip

\noindent
\begin{tabular}{@{}l p{0.8\textwidth}@{}}
Title: & \thesistitle \\
Name: & \thesisauthor \\
Date: & \thesisdate \\
Signature: & \rule{5cm}{0.4pt}
\end{tabular}


% ============================================================
% BIBLIOGRAPHY
% ============================================================
\printbibliography[heading=bibintoc]

\end{document}
